### 2019-04-10 README.Rmd
analysis -> public

---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Reassessing the Evidence for Universal School-age Bacillus Calmette Guerin (BCG) Vaccination in England and Wales

[![badge](https://img.shields.io/badge/Launch-Analysis-lightblue.svg)](https://mybinder.org/v2/gh/seabbs/AssessBCGPolicyChange/master?urlpath=rstudio)
[![Documentation](https://img.shields.io/badge/Documentation-click here!-lightgrey.svg?style=flat)](https://www.samabbott.co.uk/AssessBCGPolicyChange)
[![Paper](https://img.shields.io/badge/Paper-10.1101/567511-lightgreen.svg)](https://doi.org/10.1101/567511)
[![DOI](https://zenodo.org/badge/173767331.svg)](https://zenodo.org/badge/latestdoi/173767331)

[Sam Abbott](https://www.samabbott.co.uk), Hannah Christensen, Ellen Brooks-Pollock

## Background

In 2005, England and Wales switched from universal BCG vaccination for school-age children to targeted vaccination of neonates. We assessed some of the quantitative evidence that informed this policy change.

## Methods

We recreated a previous approach for estimating the impact of ending the BCG schools scheme in England and Wales, updating the model with parameter uncertainty. We investigated scenarios considered by the Joint Committee on Vaccination and Immunisation, and explored new approaches using notification data. We estimated the number of vaccines needed to prevent a single notification, and the average annual additional notifications caused by ending the BCG schools’ scheme.

## Results

We found a 1.9% annual decrease in Tuberculosis incidence rates best matched our estimates from notifications. We estimate that 1600 (95% Interquartile range (IQR) 1300 to 2100) vaccines would have been required to prevent a single notification in 2004. If the scheme had ended in 2001, 302 (95% IQR 238 to 369) additional annual notifications would have occurred compared to if the scheme had continued. If the scheme ended in 2016, 120 (95% IQR 88 to 155) additional annual notifications would have occurred.

## Conclusions

Our estimates of the impact of ending the BCG schools scheme were highly sensitive to the annual decrease in incidence rates. The impact of ending the BCG schools scheme was found to be greater than previously thought when parameter values were updated and notifications were used. Our results highlight the importance of including uncertainty when forecasting the impact of changes in vaccination policy.

## Reproducibility

### Repository structure

The repository is structured as an R package. It has the following structure:

- `data-raw`: Raw data processing.
- `data`: Processed data.
- `R`: Supporting R functions.
- `docs:` Documentation for R code.
- `vignettes`: Analysis paper, results, and analysis plan.
- `peer-review`: Documentation required for peer review.

### Manual install

- Install R (analysis run with `3.5.2`) and Rstudio (alternatively use Docker as outlined below).

- Download the analysis folder from [https://github.com/seabbs/AssessBCGPolicyChange/archive/master.zip](https://github.com/seabbs/AssessBCGPolicyChange/archive/master.zip) or use `git clone`, as follows, in the command line (not the R terminal).

```{bash clone-package, eval = FALSE}
git clone https://github.com/seabbs/AssessBCGPolicyChange.git
```

- Once this has been downloaded click on the project file (`AssessBCGPolicyChange.Rproj`).

- Install the analysis dependencies and build the package using the following. To enable more robust reproducibility consider using the [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/index.html) package versioned locked to R `3.5.2`.

```{r get-packages, eval = FALSE}
#install.packages("devtools")
# To build locally
devtools::install_dev_deps(dependencies = TRUE)
devtools::install()
# Alternatively to remote install
devtools::install_github("seabbs/AssessBCGPolicyChange", dependencies = TRUE)
```

- Load the analysis results by running `vignettes/paper.Rmd`. Alternatively the complete analysis (along with documentation) can be reconstructed using `make` in the project root directory.

- See `data-raw` for data processing and the documentation for implementation details.

### Docker

This analysis was developed in a docker container based on the tidyverse docker image. To run the docker image run:

```{bash, eval = FALSE}
docker run -d -p 8787:8787 --name assessbcgpolicychange -e USER=assessbcgpolicychange -e PASSWORD=assessbcgpolicychange seabbs/assessbcgpolicychange
```

The rstudio client can be found on port :8787 at your local machines ip. The default username:password is assessbcgpolicychange:assessbcgpolicychange, set the user with -e USER=username, and the password with - e PASSWORD=newpasswordhere. The default is to save the analysis files into the user directory.

If you have access to the required underlying raw data (see [`tbinenglanddataclean`](https://www.samabbott.co.uk/tbinenglanddataclean/)) then the entire analysis can be reproduced from scratch by adding the following to the `docker run` command, with the data saved into `data/tb_data`. The data requirements, and structure, can be found [here](https://www.samabbott.co.uk/tbinenglanddataclean/).

```{bash, eval = FALSE}
--mount type=bind,source=$(pwd)/data/tb_data,target=/home/DirectEffBCGPolicyChange/data/tb_data
```

Alternatively the analysis environment can be accessed via [binder](https://mybinder.org/v2/gh/seabbs/DirectEffBCGPolicyChange/master?urlpath=rstudio).

# Docker

This packge was developed in a docker container based on the tidyverse docker image. To run the docker image run:

```{bash, eval = FALSE}
docker run -d -p 8787:8787 --name assessbcgpolicychange -e USER=seabbs -e PASSWORD=seabbs seabbs/assessbcgpolicychange
```

The rstudio client can be found on port :8787 at your local machines ip. The default username:password is seabbs:seabbs, set the user with -e USER=username, and the password with - e PASSWORD=newpasswordhere. The default is to save the analysis files into the user directory.

### 2019-04-10 peer-review/cover-letter.Rmd
analysis -> public

---
title: ''
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Editor,

Please find attached our manuscript, entitled "Reassessing the Evidence for Universal School-age Bacillus Calmette Guerin (BCG) Vaccination in England and Wales".

We feel our manuscript will be of interest to you and the journal's readership because Bacillus Calmette-Guérin (BCG) is one of the most widely-used vaccines worldwide and our study explores the impact of changing BCG vaccination policy. We found that previous estimates of the impact of ending the BCG schools' scheme may have been spuriously precise as parameter uncertainty and measurement error were not fully included. Using surveillance data, we updated these estimates and found that the impact of ending the BCG schools' scheme was greater than previously estimated. These results highlight the importance of including parameter and measurement error when forecasting the impact of changes in vaccination policy. They also highlight the benefits of reproducibility as our study built on work previously published in this journal. This study benefited from access to previously published work that was used by policy makers. We combined this with more modern statistical methods to fully capture uncertainty and used recent data to increase the accuracy of the model estimates.

We do not have any issues relating to the journal policies. All authors have approved this version of the manuscript for submission and the content of the manuscript has not been published or submitted elsewhere.

With kind regards, on behalf of all the co-authors,

Sam Abbott

### 2019-04-10 R/TB_decrease_as_matrix.R

@@ -1,0 +1,29 @@
+#' Annual Decreae in TB from Single Estimate for Each Group
+#'
+#' @description From a vetor of assumed annual decreases in TB, this function constructs a
+#' matrix with yearly estimates of TB decrease
+#' @param TB_decrease A named vector of percentage decreases (i.e 1\%) in TB incidence. Names should refer to age group.
+#' @param TB_cohorts A character vector indicating the names of the cohorts modelled. Defaults to 15-19 years, 20-24 years, and 25-29 years.
+#' @param data_start Numeric, the year the first data point originates, defaults to 1969 as in Sutherland et al.
+#' @param projected_cohort Numeric, the number of cohort for which to simulate, defaults to 5. Combines with \code{cohort_length}
+#' to give the length of time to simulate.
+#' @param cohort_length Numeric, the lengh of time spent in each cohort, defaults to 5 years as in Sutherland et al.
+#' @inheritParams sutherland_model
+#' @return A matrix of annual TB incidence decreases as a percentage (0.01) for a given number of years and age groups
+#' @export
+#'
+#' @examples
+#'
+#' TB_decrease <- c(10, 8, 11)
+#'
+#'
+#' TB_decrease_as_matrix(TB_decrease)
+#'
+TB_decrease_as_matrix <- function(TB_decrease, TB_cohorts = c('15-19 years', '20-24 years', '25-29 years'), data_start = 1969, projected_cohort = 9, cohort_length = 5) {
+  TB_decrease_matrix <- t(matrix(rep(TB_decrease/100,
+                                     (projected_cohort)*cohort_length),3,(projected_cohort)*cohort_length))
+  rownames(TB_decrease_matrix) <- as.character(sapply(1:nrow(TB_decrease_matrix), function(i) { M <- data_start + i - 1 }) )
+  colnames(TB_decrease_matrix) <- names(TB_decrease)
+
+  return(TB_decrease_matrix)
+}

### 2019-04-10 R/calc_model_sum_stat.R

@@ -1,0 +1,76 @@
+
+#' Function to Calculate Summary Measures from Multiple Runs of the Sutherland et al.'s Model
+#'
+#' @param x A list of multiple simulation runs of a singl table as produced by \code{\link{AssessBCGPolicyChange}}.
+#' @param ci_level Numeric, specifies the confidence level to use. Defaults to 0.95.
+#' @param ci_space Character, string to use to seperate confidence intervals, defaults to "-"
+#' @param round_level Integer, the number of digits to which to round the summarised results. Defaults to 0
+#'
+#' @return A list of tables for a set of summary measures. These are; mean with CI, median, 2.5% quantile,
+#' 97.5% quantile, 25%  quantile, 75% quantile, mean, lower CI, and upper CI.
+#' @export
+#'
+#' @import magrittr
+#' @importFrom purrr map
+#' @examples
+#'
+#'## Run multiple model simulations
+#'model_sims <- simulate_sutherland_model(cores = 1)
+#'
+#'## Summarise the Total effects of ending the scheme table
+#'calc_model_sum_stat(model_sims[["Table 5 - Total Effects of ending the schools BCG scheme"]])
+
+calc_model_sum_stat <- function(x, ci_level = 0.95, ci_space = "-", round_level = 0) {
+
+  CI <- function(x,ci = .95) {
+    a<-mean(x)
+    s<-sd(x)
+    n<-length(x)
+    error<-qt(ci+(1-ci)/2,df=n-1)*s/sqrt(n)
+      return(c(upper = a + error,mean = a,lower = a - error))
+    }
+
+  ## convert objects to matrices
+  if(!is.matrix(x[[1]]))
+  {
+    x <- x %>% map(as.matrix)
+  }
+
+  ##Preallocate
+  sim_results_mean_CI <- sim_results_median <- sim_results_mean <- sim_results_lower <- sim_results_upper <- matrix(NA, nrow=nrow(x[[1]]), ncol=ncol(x[[1]]))
+  sim_results_med_IQR <- sim_results_2.5 <- sim_results_97.5 <- sim_results_25 <- sim_results_75 <- sim_results_mean_CI
+  samples <- length(x)
+
+  for(i in 1:nrow(x[[1]]))
+  {
+    for(j in 1:ncol(x[[1]]))
+    {
+      tmp <- vector(length=samples)
[…]

### 2019-04-10 R/fit_model_to_data.R

@@ -1,0 +1,34 @@
+#' Fit Percentage of First Year Cases
+#'
+#' @description Fit the percentage of cases in the first year using the estimation model and
+#' original results of sutherland et al. using sum of least squares.
+#'
+#' @param model Model function to fit, by default uses \code{\link[AssessBCGPolicyChange]{sutherland_model}}.
+#' @param results Orginal results, by default uses \code{\link[AssessBCGPolicyChange]{sutherland_results}}
+#' @param ... Pass additional to \code{model}. See \code{\link[AssessBCGPolicyChange]{sutherland_model}} for details if
+#' using this as the \code{model} function.
+#'
+#' @return The best fitting value for the percentage of cases in the first year using the sum of least squares.
+#' @export
+#'
+#' @examples
+#'
+#'fit_model_to_data(model = sutherland_model, results = sutherland_results)
+#'
+fit_model_to_data <- function(model = sutherland_model, results = sutherland_results, ...){
+
+
+   model_fit <- optimize(function(x) {
+
+    model_results <- model(Percentage.Year.One = x, ...)
+    model_total_nots <- model_results[["Table 5 - Total Effects of ending the schools BCG scheme"]][c('1986','1991','1996'), c('1988', '1993', '1998', '2003', '2008', '2013')]
+
+    sum_of_squares <- sum((results - model_total_nots)^2)
+
+    return(sum_of_squares)
+     },
+    interval = c(0,1)
+   )
+
+return(model_fit$minimum)
+  }

### 2019-04-10 R/gen_annual_change_scenarios.R

@@ -1,0 +1,36 @@
+#' Function to Generate Annual Change Scenarios Based on Sutherland Assumptions
+#'
+#' @param TB_decrease A numeric vector of percentage decreases in TB notifications rates (formated as 0.09 etc.).
+#' @param cohorts_to_sim The number of cohorts to simulate. Defaults to 11
+#' @param samples Numeric, the number of samples to generate for each scenario. Defaults to 10
+#'
+#' @return A dataframe with each row representing a sample, with annual decrease stored as a matrix
+#' @export
+#'
+#' @importFrom tibble tibble
+#' @importFrom purrr map transpose flatten
+#' @seealso TB_decrease_as_matrix
+#' @examples
+#'
+#' gen_annual_change_scenarios()
+gen_annual_change_scenarios <- function(TB_decrease = c(0.09, 0.05, 0.01),
+                                        cohorts_to_sim = 11, samples = 10){
+
+
+  TB_decrease_list <- map(1:samples, function(i) {
+    map(TB_decrease, ~TB_decrease_as_matrix(rep(. * 100, 3),
+                        TB_cohorts = c("15-19 years", "20-24 years", "25-29 years"),
+                        data_start = 1969,
+                        projected_cohort = cohorts_to_sim,
+                        cohort_length = 5)
+        )
+  }) %>% transpose
+
+ TB_decrease_name <- paste0(round(TB_decrease * 100, digits = 1), "% decrease")
+
+ scenarios <- tibble(scenario = map(TB_decrease_name, ~rep(., samples)) %>% unlist,
+                     sample = rep(1:samples, length(TB_decrease_name)),
+                     annual_TB_decrease = flatten(TB_decrease_list))
+
+ return(scenarios)
+}

### 2019-04-10 R/gen_scenario_results.R

@@ -1,0 +1,55 @@
+#' A Function to Generate Scenario Results from the Sutherland Model
+#'
+#' @param scenarios A dataframe of scenarios, see the examples for a structural example.
+#' @param parameter_samples A dataframe of parameter samples, see the examples for an example.
+#' @param cores The number of cores to use, defaults to one
+#'
+#' @return A dataframe combining the scenarios, parameter samples, and model results
+#' @export
+#' @import magrittr
+#' @importFrom dplyr full_join bind_cols select
+#' @importFrom purrr flatten map
+#' @importFrom tibble tibble
+#' @importFrom tidyr spread
+#' @examples
+#'
+#' scenarios <- gen_annual_change_scenarios(TB_decrease = c(0.09, 0.05, 0.01),
+#' cohorts_to_sim = 11, samples = 10)
+#'
+#' parameter_samples <- sample_model_parameters(incidence_rates = sutherland_incidence_rates,
+#' population_data = sutherland_data, samples = 10)
+#'
+#' gen_scenario_results(scenarios, parameter_samples, cores = 1)
+gen_scenario_results <- function(scenarios, parameter_samples, cores = 1) {
+
+  if (max(scenarios$sample) != max(parameter_samples$sample)) {
+    stop("The number of samples in scenarios and parameter_samples must match")
+  }
+
+  ## Bind scenarios and parameter samples
+  scenarios <- scenarios %>%
+    full_join(parameter_samples, by = "sample")
+
+  ## Simulate model - map results into data frame
+  model_sims <- simulate_sutherland_model(list_annual_TB_decrease = scenarios$annual_TB_decrease,
+                                                          list_incidence_rates = scenarios$incidence_rates,
+                                                          list_percentage_year_one = scenarios$per_year_one,
+                                                          list_sym_lag = scenarios$sym_lag,
+                                                          cores = cores)
+
+  model_sims <- tibble(output =  map(names(model_sims),
+                                     ~rep(., length(flatten(model_sims))/length(names(model_sims)))) %>%
+                         unlist,
+                       simulation = model_sims %>% flatten,
+                       sample = rep(1:as.integer(length(flatten(model_sims))/length(names(model_sims))),
+                                    length(names(model_sims)))
+  ) %>%
+    spread(output, simulation)
+
+  ## Join scenarios and samples
+  scenario_results <- scenarios %>%
+    bind_cols(model_sims %>%
+                select(-sample))
+
+  return(scenario_results)
+}

### 2019-04-10 R/inc_rates_uk_e.R

@@ -1,0 +1,17 @@
+#' Incidence Rates per 100,000 in the UK born in England
+#'
+#' A dataset containing the estimate incidence rates and associated variables for the UK born in England. Estimated using the Labour Force Survey and the Enhanced Tubercuolosis
+#' Surveillance system. See [here](https://www.samabbott.co.uk/tbinenglanddataclean)  for details.
+#' @format A dataframe with 51 rows and 6 variables
+#'
+#' \describe{
+#'   \item{year}{The calendar year}
+#'   \item{age_group}{Age groups used in the model, these are 15-19 years old, 20-24 years old and 25-29 years old.}
+#'   \item{incidence}{The estimated incidence per 100,000 individuals}
+#'   \item{lci}{The lower confidence interval \(95% CI\)}
+#'   \item{uci}{The lower confidence interval \(95% CI\)}
+#'   \item{sd}{The standard deviation of the estimated incidence rates}
+#'   }
+#'
+#'
+"inc_rates_uk_e"

### 2019-04-10 R/nots_ew.R

@@ -1,0 +1,12 @@
+#' Tuberculosis Notifications in England and Wales 1913-2016
+#'
+#' A dataset containing the number of notifications of Tuberculosis, in any given year, in England
+#' and Wales. Originally published by Public Health England, see [here](https://www.samabbott.co.uk/tbinenglanddataclean/reference/tb_not_ew.html) for details.
+#' @format A dataframe with 104 rows and 2 columns
+#' \describe{
+#'   \item{year}{The calendar year}
+#'   \item{notifications}{The number of notified cases in any given year.}
+#' }
+#'
+#'
+"nots_ew"

### 2019-04-10 R/sample_annual_decrease_inc_rates.R

@@ -1,0 +1,198 @@
+
+#' Samples the Percentage Annual Decrease Based On Incidence Rates
+#'
+#' @description This function samples the percentage annual decrease, using Sutherland et al's estimates for
+#' White UK born incidence rates and estimates of UK born incidence rates  in England calculated using the Enhanced Tuberculosis
+#' surveilliance system and the Labour Force Survey. Incidence rates are first sampled from the estimated UK born incidence rates,
+#' by assuming that they are normally distributed. Incidence rates are then sampled from the Sutherland et al. estimates from 1973, 1978, and 1983
+#' using \code{\link[AssessBCGPolicyChange]{sample_incidence_rates}}. Average incidence rates for all cases are then estimated using Sutherland
+#' et al.'s population weights for those BCG vaccinated (75%), unvaccinated (11%), tuberculin positive (8%) and ineligible (6%). Missing
+#' incidence rates are then estimated using a loess fit for each sample, with samples predicted to have an incidence rate below 0.1 per 100,000
+#' having this set to be 0.1 per 100,000. The percentage annual difference is the calculated for each sample
+#'  using  \code{\link[AssessBCGPolicyChange]{ sample_per_decrease}}. Projected annual percentage changes are estimated using an average of
+#'  the previous 3 years. Similarly for years prior to Sutherland et als incidence rate estimates an average of the previous 3 years are used.
+#'  Samples of the percentage annual decrease are returned as a tibble, containing the nested dataframe and a nested matrix ready to be used by
+#'  \code{\link[AssessBCGPolicyChange]{sutherland_model}}.
+#' @param samples Numeric, the number of samples to take. Defaults to 10.
+#' @param verbose Logical, defaults to \code{TRUE}. Should summary measures and plots be returned.
+#' @return A nested tibble, containing the sample number, and both a nested dataframe and nested matrix of
+#' annual percentage changes.
+#' @seealso sample_per_decrease sample_annual_decrease_nots
+#' @inheritParams sample_annual_decrease_nots
+#' @export
+#' @import magrittr
+#' @import ggplot2
+#' @import multidplyr
+#' @import dplyr
[…]

### 2019-04-10 R/sample_annual_decrease_nots.R

@@ -1,0 +1,63 @@
+
+#' Samples the Percentage Annual Decrease Based On Notifications
+#'
+#' @description This function samples the percentage annual decrease, using TB notifications.
+#' The percentage annual change is first estimated using \code{prop.test}, along with the 95% confidence
+#' intervals. Assuming that the percentage annual decrease is normally distributed multiple samples are then
+#' taken. This is returned as a tibble, containing the nested dataframe and a nested matrix ready to be
+#' used by \code{\link[AssessBCGPolicyChange]{sutherland_model}}.
+#' @param notifications A dataframe of notifications, it must contain a year variable, and the second variable
+#' must be the number of notifications. If not specified defaults to \code{\link[AssessBCGPolicyChange]{nots_ew}}.
+#' @param samples Numeric, the number of samples to take. Defaults to 100.
+#' @return A nested tibble, containing the sample number, and both a nested dataframe and nested matrix of
+#' annual percentage changes.
+#' @seealso sample_per_decrease
+#' @inheritParams TB_decrease_as_matrix
+#' @inheritParams sample_per_decrease
+#' @export
+#' @import magrittr
+#' @importFrom dplyr filter mutate left_join select group_by ungroup
+#' @importFrom tibble tibble
+#' @importFrom tidyr spread nest
+#' @importFrom purrr map
+#' @examples
+#'
+#' sample_annual_decrease_nots()
+sample_annual_decrease_nots <- function(notifications = NULL,
+                                        TB_cohorts = c("15-19 years", "20-24 years", "25-29 years"),
+                                        samples = 100, max_year = NULL, years_to_avg = 3) {
+  if(is.null(notifications)) {
+    nots <- nots_ew
+  }else{
+    nots <- notifications
+  }
+
+  ## Sample percentage decrease of TB
+  nots <- nots_ew %>%
+    filter(year >= 1968) %>%
+    sample_per_decrease(samples = samples, max_year = max_year,
+                        years_to_avg = years_to_avg)
+
+  ## Bind percentage decrease to age groups
+  age_groups <- tibble(age_groups = TB_cohorts, key = 1)
+
+  nots_by_age <- nots %>%
+    mutate(key = 1) %>%
+    left_join(age_groups, by = "key") %>%
+    select(year, sample, age_groups, prop_decrease_sample) %>%
[…]

### 2019-04-10 R/sample_inc_rate.R

@@ -1,0 +1,41 @@
+
+#' Function to Sample from the Underlying Distribution of an Incidence Rate
+#'
+#' @description The number of cases are back calculated using the incidence rate and
+#' the size of the population. Samples are then taken of notifications, assuming a poisson
+#' distribution. Incidence rates are then calculated based on these samples. Multiple samples
+#' of a matrix input may be taken with the output formated as a list.
+#' @param inc_rate Incidence rates formated as \code{\link[AssessBCGPolicyChange]{sutherland_incidence_rates}}.
+#' @param pop Population estimates for each cohort as a vector.
+#' @param inc_per The scaling factor for the incidence rate, defaults to 100,000.
+#' @param samples Numeric, the number of samples to take. Defaults to 100.
+#'
+#' @return Incidence rate samples as a list, formated as a 3 by 3 matrix.
+#' @export
+#'
+#' @importFrom purrr map_dbl map
+#' @examples
+#'
+#' sample_inc_rate(Unvaccinated, sutherland_data$BCG.Unvaccinated[1:3], samples = 5)
+#'
+sample_inc_rate <- function(inc_rate, pop, inc_per = 100000, samples = 100) {
+
+    nots <- inc_rate * pop / inc_per
+    nots_flat <- as.vector(nots)
+    pop <- as.vector(pop)
+
+    sample_nots <- map(1:samples, function(i){
+      sample_nots <- purrr::map_dbl(nots_flat, ~rpois(n = 1, lambda = .))
+      sample_inc_rate <- sample_nots / pop * inc_per
+
+      sample_inc_rate <- matrix(sample_inc_rate, 3, 3)
+      colnames(sample_inc_rate) <- colnames(inc_rate)
+      rownames(sample_inc_rate) <- rownames(inc_rate)
+
+
+      return(sample_inc_rate)
+    })
+
+    return(sample_nots)
+}
+

### 2019-04-10 R/sample_incidence_rates.R

@@ -1,0 +1,39 @@
+#' Function to sample Sutherland Incidence rates to Include Uncertainty
+#'
+#' @param incidence_rates Incidence rates as formated by \code{\link[AssessBCGPolicyChange]{sutherland_incidence_rates}}
+#' @param population_data Population data as formated by  \code{\link[AssessBCGPolicyChange]{sutherland_data}}
+#' @param samples Numeric, the number of samples to take. Defaults to 100.
+#'
+#' @return A list of lists. Each list contains matrices of sampled incidence rates for unvaccinated, vaccinated, tuberculin postive
+#' and inelgible populations.
+#' @export
+#' @importFrom purrr map2 transpose
+#' @import magrittr
+#' @examples
+#'
+#' sample_incidence_rates(sutherland_incidence_rates, sutherland_data, samples = 5)
+#'
+sample_incidence_rates <- function(incidence_rates, population_data, samples = 100) {
+  ## Sample incidence rates to correctly include uncertainty
+  unvac_inc_sample <- sample_inc_rate(incidence_rates$Unvaccinated,
+                                      population_data$BCG.Unvaccinated[1:3], samples = samples)
+  vac_inc_sample <- sample_inc_rate(incidence_rates$Vaccinated,
+                                    population_data$BCG.vaccinations[1:3], samples = samples)
+
+  not_prevented <- map2(unvac_inc_sample, vac_inc_sample, function(x, y) x-y)
+
+  tub_inc_sample <- sample_inc_rate(incidence_rates$Tuberculin,
+                                    population_data$BCG.Tuberculin[1:3], samples = samples)
+
+  ## Assumption that those inelgible have the same incidence rates as the unvaccinated. No published data avaialble
+  inelg_inc_sample <- unvac_inc_sample
+
+  incidence_rates <- list(unvac_inc_sample, vac_inc_sample, not_prevented, tub_inc_sample, inelg_inc_sample)
+  names(incidence_rates) <- c("Unvaccinated", "Vaccinated", "Notifications.Prevented", "Tuberculin", "Inelgible")
+
+  incidence_rates <- incidence_rates %>%
+    transpose
+
+  return(incidence_rates)
+
+}

### 2019-04-10 R/sample_model_parameters.R

@@ -1,0 +1,51 @@
+#' Function to Sample Model Parameters
+#'
+#' @description The serial interval is estimated as 1.44 (95% CI 1.29 to 1.63) years based on
+#' [this](https://doi.org/10.1093/ije/dyr058) study. Samples were taken by assuming a normal
+#' distribution with a standard deviation estimated from the parameters confidence intervals.
+#' The above study includes data on the percentage of cases in each year since infection in the
+#' form of a figure. This appears to roughly match the original model estimate. As no other estimate
+#' is available the point estimate from the original model is used.
+#' @inheritParams sample_incidence_rates
+#' @param samples Numeric, the number of samples to run. Defaults to 100
+#' @param as_sutherland Logical, defaults to \code{FALSE}. If \code{TRUE} then the
+#' original estimates used by sutherland et al. are returned. Note these are point estimates.
+#' @return A tibble of parameter samples
+#' @export
+#' @seealso sample_incidence_rates
+#' @importFrom tibble tibble
+#' @examples
+#'
+#' sample_model_parameters()
+#'
+sample_model_parameters <- function(incidence_rates = sutherland_incidence_rates,
+                                    population_data = sutherland_data,
+                                    samples = 100, as_sutherland = FALSE){
+
+  if (as_sutherland) {
+
+    per_year_one <- 0.764
+    sym_lag <- 2
+    inc_rates <- list(sutherland_incidence_rates)
+
+  }else{
+    ## Generate parameter estimates
+    per_year_one <- 0.764
+
+    point_sym_lag <- 1.44
+    sd_sym_lag <-  (1.63 - 1.29)/(2*qnorm(0.975))
+
+    sym_lag <- rnorm(samples,  point_sym_lag, sd_sym_lag)
+
+    inc_rates <- sample_incidence_rates(incidence_rates = sutherland_incidence_rates,
+                                        population_data = sutherland_data,
+                                        samples = samples)
+  }
+
+  sample_parameters <- tibble(sample = 1:samples,
+                              incidence_rates = inc_rates,
+                              per_year_one = per_year_one,
+                              sym_lag = sym_lag)
+
+  return(sample_parameters)
+}

### 2019-04-10 R/sample_per_decrease.R

@@ -1,0 +1,94 @@
+#' A Function to Sample the Annual Percentage Decrease with Uncertainty
+#'
+#' @param df A dataframe containing a year variable and with the second variable being the
+#' one to sample from.
+#' @param samples Numeric, the number of samples to take.
+#' @param max_year Numeric, set as greater than the maximum year in the dataframe if interested in projection.
+#' Defaults to \code{NULL}. If specified, then an average of the percentage decrease for a set number
+#' of years is taken. This is specified with \code{years_to_avg}.
+#' @param min_year Numeric, set as smaller than the minimum year in the dataframe if interested in backwards projection.
+#' Defaults to \code{NULL}. If specified, then an average of the percentage decrease for a set number
+#' of years is taken. This is specified with \code{years_to_avg}.
+#' @param years_to_avg Numeric, the number of years over which to average the percentage decrease
+#' when estimating the ongoing percentage decrease.
+#' @return A dataframe containing annual percentage decrease samples.
+#' @export
+#' @import magrittr
+#' @importFrom dplyr mutate select full_join everything summarise_all ungroup
+#' @importFrom tidyr unnest
+#' @importFrom purrr map2_dbl map2
+#' @importFrom tibble tibble
+#' @examples
+#'
+#'  df <- data.frame(year = c(2000:2010), measure = seq(10, 110, 10))
+#'
+#' sample_per_decrease(df, samples = 2)
+#'
+sample_per_decrease <- function(df, max_year = NULL, min_year = NULL, samples = 100, years_to_avg = 3) {
+  df$measure <- df[[2]]
+
+  ## Internal function to estimate the standard devation using prop.test
+  prop_sd <- function(p, n) {
+    if (p > n) {
+      num <- n
+      denom <- p
+    }else{
+      num <- p
+      denom <- n
+    }
+    prop <- suppressWarnings(prop.test(num, denom))
+
+    ## Estimate ci and back calc sd
+    lci <- prop$conf.int[1]
+    uci <- prop$conf.int[2]
+    sd <- (uci-lci)/(2*qnorm(0.975))
+
+    return(sd)
+  }
+
+  df_sample <- df %>%
+    mutate(prev_year_m = lag(measure)) %>%
+    mutate(prop_decrease =  1 - measure / prev_year_m) %>%
+    na.omit()
+
[…]

### 2019-04-10 R/simulate_sutherland_model.R

@@ -1,0 +1,49 @@
+#' Function to run multiple simulations of the Sutherland et al. Estimation Model
+#' @param list_annual_TB_decrease A list of annual TB decreases with each element as specified for
+#' the \code{Annual.TB.Decrease.Yearly} arguement of \code{\link[AssessBCGPolicyChange]{sutherland_model}}.
+#' @param list_percentage_year_one A list of percentage secondary notifications in the first year
+#'  with each element as specified for the \code{Percentage.Year.One} arguement of
+#'  \code{\link[AssessBCGPolicyChange]{sutherland_model}}.
+#' @param list_incidence_rates A list of incidence rates with each element as specified for the \code{incidence_rates}
+#'  arguement of \code{\link[AssessBCGPolicyChange]{sutherland_model}}.
+#' @param cores The number of cores to use, defaults to one
+#' @param ... Additional parameters to pass to \code{\link[AssessBCGPolicyChange]{sutherland_model}}.
+#'
+#' @return A list of simulations for each of the outputs of \code{\link[AssessBCGPolicyChange]{sutherland_model}}.
+#' @export
+#' @importFrom purrr transpose
+#' @importFrom doParallel registerDoParallel stopImplicitCluster
+#' @import foreach
+#'
+#' @examples
+#'
+#' simulate_sutherland_model(cores = 1)
+#'
+simulate_sutherland_model <- function(list_annual_TB_decrease = list(TB_decrease_as_matrix(sutherland_TB_decrease),
+                                                                     TB_decrease_as_matrix(sutherland_TB_decrease)),
+                                      list_incidence_rates = list(sutherland_incidence_rates, sutherland_incidence_rates),
+                                      list_percentage_year_one = rep(0.764, 2), list_sym_lag = list(2, 2),
+                                      cores = 1, ...) {
+  if (cores > 1) {
+    doParallel::registerDoParallel(cores = cores)
+    }
+
+
+  samples <- length(list_annual_TB_decrease)
+
+  simulation_estimates <- foreach(i = 1:samples) %dopar%
+  {
+    sutherland_model(incidence_rates = list_incidence_rates[[i]],
+                     Annual.TB.Decrease.Yearly = list_annual_TB_decrease[[i]],
+                     Percentage.Year.One = list_percentage_year_one[[i]],
+                     Sym.Lag = list_sym_lag[[i]], ...)
+  }
+
+  if (cores > 1) {
[…]

### 2019-04-10 R/sutherland_TB_decrease.R

@@ -1,0 +1,7 @@
+#' Sutherland Assumption of Annual TB decrease
+#'
+#' A dataset containing the estimated annual decrease in TB notifications from Sutherland et al.'s original paper. See
+#'  \code{sutherland_TB_decrease.R} (see \code{data-raw}) and the vignette discussing recreating Sutherland et al.'s model.
+#' @format A vector with 3 elements. Each element represents an age group (15-19 years 20-24 years 25-29 years) and contains the projected
+#' decrease in TB for that age group.
+"sutherland_TB_decrease"

### 2019-04-10 R/sutherland_data.R

@@ -1,0 +1,23 @@
+#' Data used in Sutherland et al
+#'
+#' A dataset containing the data used by Sutherland et al to estimate the impact of ending the BCG school's
+#' scheme. The data has been aggregated for surveys. Details of its collection can be found in the original paper,
+#' in \code{sutherland_data.R} (see \code{data-raw}) and in the vignette discussing recreating Sutherland et al.'s model.
+#' @format A list with 14 elements.
+#' \describe{
+#'   \item{Cohort}{A character string indicating the year of entry to study}
+#'   \item{Cohort.Ranges}{Age ranges included in the study; 15-19, 20-24 and '25-29 years}
+#'   \item{Projected.Cohort}{The number of years for which to project forward (scaled by the length of cohort)}
+#'   \item{Cohort.Pop}{The population at entry in each cohort}
+#'   \item{Unvaccinated}{Incidence rates (per 100,000) in those unvaccinated by cohort and age group}
+#'   \item{Vaccinated}{Incidence rates (per 100,000) in those vaccinated by cohort and age group}
+#'   \item{Notifications.Prevented}{Notifications prevented by 100,000 people by cohor and age groups}
+#'   \item{BCG.efficacy}{Estimated efficacy of the BCG vaccine}
+#'   \item{BCG.vaccinations}{Number of BCG vaccines given by 5 year period}
+#'   \item{BCG.Unvaccinated}{Number of BCG unvaccinated by 5 year period}
+#'   \item{BCG.Tuberculin}{Number Tuberculin positive by 5 year period}
+#'   \item{BCG.Ineglible}{Number inelgible for the BCG schools scheme by 5 year period}
+#' }
+#'
+#'
+"sutherland_data"

### 2019-04-10 R/sutherland_gen_time.R

@@ -1,0 +1,9 @@
+#' Assumed Generation time From Sutherland et al.
+#'
+#' A dataset containing the assumed generation time from Sutherland et al.'s original paper. See
+#'  \code{sutherland_gen_time.R} (see \code{data-raw}) and the vignette discussing recreating Sutherland et al.'s model.
+#' @format A dataframe with 1 variable and 1 row
+#' \describe{
+#' \item{sutherland_gen_time}{The assumed generation time}
+#' }
+"sutherland_gen_time"

### 2019-04-10 R/sutherland_incidence_rates.R

@@ -1,0 +1,15 @@
+#' Incidence rates used in Sutherland et al
+#'
+#' A dataset containing the incidence rates used by Sutherland et al to estimate the impact of ending the BCG school's
+#' scheme. The data has been aggregated for surveys. Details of its collection can be found in the original paper,
+#' in \code{sutherland_incidence_rates.R} (see \code{data-raw}) and in the vignette discussing recreating Sutherland et al.'s model.
+#' @format A list with 14 elements.
+#' \describe{
+#'   \item{Unvaccinated}{A matrix of incidence rates for the unvaccinated by age and year of entry}
+#'   \item{Vaccinated}{A matrix of incidence rates for the vaccinated by age and year of entry}
+#'   \item{Notifications.Prevented}{A matrix of notifications prevented (as rates) by age and year of entry}
+#'   \item{Tuberculin}{Incidence rates (per 100,000) in the Tuberculin positive population}
+#'   \item{Inelgible}{Incidence rates (per 100,000) in those inelgible for BCG vaccination}
+#'}
+#'
+"sutherland_incidence_rates"

### 2019-04-10 R/sutherland_model.R

@@ -1,0 +1,312 @@
+
+#' Sutherland et al. Model to Estimate Impact of Ending the BCG School Scheme
+#'
+#' @param Data Data used by Sutherland et al. analysis. See \code{\link[AssessBCGPolicyChange]{sutherland_data}}.
+#' @param incidence_rates A list of Incidence rates (per 100,000) for the BCG vaccinated and unvaccinated population.
+#' See \code{\link[AssessBCGPolicyChange]{sutherland_incidence_rates}} for details.
+#' @param Rates.Per The number of people by which to produce rates, defaults to 100,000 as in Sutherland et al.
+#' @param Cohort.Length Numeric, the lengh of time spent in each cohort, defaults to 5 years as in Sutherland et al.
+#' @param Data.start Numeric, the year the first data point originates, defaults to 1969 as in Sutherland et al.
+#' @param Annual.TB.Decrease.Yearly Matrix of average yearly decreases in TB with each column representing an age group (15-19, 20-24, 25-29).
+#' Row names must be the year, starting from \code{Data.start}. Defaults to using  \code{\link[AssessBCGPolicyChange]{sutherland_TB_decrease}}.
+#' @param Percentage.Year.One Numeric, the percentage of a generation that occur in the first year.
+#' Fitted via, \code{\link[AssessBCGPolicyChange]{fit_model_to_data}}.
+#' @param trans_params A funciton for estimating the transmission parameters (expected total secondary notifications,
+#' the size of the first generation, and the average interval between a primary case and all secondary cases). Defailts to
+#' \code{\link[AssessBCGPolicyChange]{sutherland_trans_params}}.
+#' @param Sym.Lag Numeric, the generation time between infection and symtoms. Must be smaller or equal to the \code{Cohort.Length}.
+#' Defaults to \code{\link[AssessBCGPolicyChange]{sutherland_gen_time}}.
[…]

### 2019-04-10 R/sutherland_results.R

@@ -1,0 +1,7 @@
+#' Results for Total notifications From Sutherland et al.
+#'
+#' A dataset containing the estimated number of notifications for Sutherland et al.'s original paper. See
+#'  \code{sutherland_results.R} (see \code{data-raw}) and the vignette discussing recreating Sutherland et al.'s model.
+#' @format A matrix with 3 rows and 6 columns. The column names indicate the estimated number of notifications in a given year,
+#' and the rows indicate which year the schools scheme was ended.
+"sutherland_results"

### 2019-04-10 R/sutherland_trans_params.R

@@ -1,0 +1,41 @@
+#' Estimate Transmission chain parameters using Sutherland et al.s Approach
+#'
+#' @param Sutherland Logical, defaults to \code{TRUE}. Whether or not to use Sutherland et al.'s
+#' assumption for estimating the basic reproduction number.
+#' @param Repro.No If \code{Sutherland} is \code{FALSE} sets the basic reproduction number used to estimate
+#' the transmission chain parameters.
+#' @return A dlist containing the expected total secondary notifications (Expected.Total.Sec.Note),
+#' the size of the first generation (Size.First.Gen) and the average interval to all secondary cases (Avg.Int.All.Sec).
+#' @export
+#' @inheritParams sutherland_model
+#' @examples
+#'
+#'sutherland_trans_params()
+#'
+sutherland_trans_params <- function(Annual.TB.Decrease.Yearly = TB_decrease_as_matrix(sutherland_TB_decrease), Sym.Lag = sutherland_gen_time,
+                                    Sutherland = TRUE, Repro.No = NULL,
+                                    Cohort.Length = 5)
+{
+
+  ################################## Calculate parameters for secondary notifcations model
+  ################ Expected Total no. notifications
+  if (Sutherland) {
+    Expected.Total.Sec.Note <- (1 - rowMeans(Annual.TB.Decrease.Yearly))^Sym.Lag
+  }else {
+    Expected.Total.Sec.Note <- rep(Repro.No, nrow(Annual.TB.Decrease.Yearly))
+    names(Expected.Total.Sec.Note) <- rownames(Annual.TB.Decrease.Yearly)
+  }
+
+
+  ############### Size of first Gen
+  Size.First.Gen <- Expected.Total.Sec.Note / (Expected.Total.Sec.Note + 1)
+  names(Size.First.Gen) <- rownames(Annual.TB.Decrease.Yearly)
+  ############### average interval to all secondary notifications
+
+  Avg.Int.All.Sec <- Sym.Lag/(1 - Size.First.Gen)
+  names(Avg.Int.All.Sec) <- rownames(Annual.TB.Decrease.Yearly)
+
+  out <- list(Expected.Total.Sec.Note, Size.First.Gen, Avg.Int.All.Sec)
+  names(out) <- c("Expected.Total.Sec.Note", "Size.First.Gen", "Avg.Int.All.Sec")
+  return(out)
+}

### 2019-04-10 data-raw/inc_rates_uk_e.R

@@ -1,0 +1,45 @@
+library(tbinenglanddataclean)
+library(tidyverse)
+library(epiR)
+
+## Calculate incidence rates using the tbinenglandataclean - see ?tbinenglanddataclean for details
+# See: https://www.samabbott.co.uk/tbinenglanddataclean/
+inc_rates_list <- calculate_incidence_ets_lfs_ons(data_path = "~/data-raw/tbinenglanddataclean",
+                                                ets_name = "clean_ets_2016.rds",
+                                                demo_name = "E_ons_lfs_2000_2016.rds",
+                                                return = TRUE,
+                                                save = FALSE,
+                                                incidence_name = "incidence" ,
+                                                grouped_incidence_name = "age_grouped_incidence",
+                                                condensed_grouped_incidence_name = "condensed_age_group_incidence",
+                                                cases_demo_incidence_name = "cases_demo_incidence")
+
+
+## Extract just grouped incidence rates for target age groups, UK born
+inc_rates_uk_e <- inc_rates_list[[2]] %>%
+  filter(`Age group` %in% c("15-19", "20-24", "25-29")) %>%
+  filter(CoB %in% "UK born") %>%
+  select(-CoB) %>%
+  rename(age_group = `Age group`,
+         incidence = Incidence,
+         lci = Inc_LCI,
+         uci = Inc_UCI,
+         year = Year) %>%
+  mutate(year = as.numeric(year)) %>%
+  mutate(age_group = as.character(age_group) %>%
+           paste0(" years") %>%
+           factor) %>%
+  filter(year != max(year))
+
+
+## Add standard deviation assuming that it is normal and that CI's are 0.95
+inc_rates_uk_e <- inc_rates_uk_e %>%
+  ungroup %>%
+  mutate(sd =  (uci-lci)/(2*qnorm(0.975)))
+
+
+#' Add to package and keep in data-raw
+devtools::use_data(inc_rates_uk_e, overwrite = TRUE)
+write_csv(inc_rates_uk_e, "inc_rates_uk_e.csv")
+
+

### 2019-04-10 data-raw/nots_ew.R

@@ -1,0 +1,18 @@
+
+library(tidyverse)
+## Load in the data from tbinenglanddataclean
+## see www.samabbott.co.uk/tbinenglanddataclean for details
+
+nots_ew <- tbinenglanddataclean::tb_not_ew
+
+
+## Total nots repiratory/non-respiratory and pulmonary/non-pulmonary
+nots_ew <- nots_ew %>%
+  gather(key = "tb_type", value = "notifications", -year) %>%
+  group_by(year) %>%
+  summarise(notifications = sum(notifications, na.rm = T))
+
+
+#' Add to package and keep in data-raw
+devtools::use_data(nots_ew, overwrite = TRUE)
+write_csv(nots_ew, "nots_ew.csv")

### 2019-04-10 data-raw/package_set_up.R

@@ -1,0 +1,39 @@
+#install.packages("devtools")
+library(devtools)
+
+# init stuffs
+use_data_raw()
+use_cran_badge()
+use_mit_license()
+use_readme_rmd()
+use_testthat()
+use_travis()
+use_roxygen_md()
+
+
+#Suggest
+use_package("pkgdown", type = "Suggests")
+use_package("devtools", type = "Suggests")
+use_package("pomp", type = "Suggests")
+use_package("multidplyr", type = "Suggests")
+#Import
+use_package("tidyverse", type = "Imports")
+use_package("tbinenglanddataclean", type = "Imports")
+use_package("prettypublisher", type = "Imports")
+use_package("ggplot2", type = "Imports")
+use_package("doParallel", type = "Imports")
+use_package("foreach", type = "Imports")
+use_package("purrr", type = "Imports")
+
+#Vignettes
+use_vignette("analysis_plan")
+use_vignette("recreate_sutherland")
+use_vignette("update_sutherland")
+use_vignette("paper")
+use_vignette("si")
+
+##Build site, and make pkgdown
+devtools::document()
+devtools::build_vignettes()
+devtools::build()
+pkgdown::build_site()

### 2019-04-10 data-raw/sutherland_TB_decrease.R

@@ -1,0 +1,8 @@
+## Sutherland assumed annual TB decrease in each age group
+sutherland_TB_decrease <- c(9,9,9)
+
+devtools::use_data(sutherland_TB_decrease, overwrite = TRUE)
+readr::write_csv(as.data.frame(sutherland_TB_decrease), "sutherland_TB_decrease.csv")
+
+
+

### 2019-04-10 data-raw/sutherland_data.R

@@ -1,0 +1,45 @@
+
+##################### Build 1983 data
+Cohort <- c('1967-71','1972-1976','1977-1981')
+Cohort.Ranges <- c('15-19 years', '20-24 years', '25-29 years')
+
+##################### Population of Cohorts
+Cohort.Pop <- t(c(2970000,3448000,3666000))
+colnames(Cohort.Pop) <- Cohort
+
+################### Estimated Efficacy of vaccine per 100,000
+BCG.efficacy <- c(80,73,67)
+names(BCG.efficacy) <- Cohort.Ranges
+
+################## No. of BCG vaccinations given:
+
+BCG.vaccinations <- c(2.210,2.548,2.892,2.55,2.00,2.10)  #Whilst Sutherland states these are 75% of populations, for years surveyed they are actual data.
+BCG.vaccinations <- BCG.vaccinations*10^6
+names(BCG.vaccinations) <- c('1967-71','1972-1976','1977-1981','1982-86','1987-91','1992-1996')
+
+################# No. of Negative unvaccinated people: set to be the average of collected data at 11%
+BCG.Unvaccinated <- c(0.465*0.95, 0.664 * 0.943, 0.585 * 0.936)
+BCG.Unvaccinated <- BCG.Unvaccinated*10^6
+BCG.Unvaccinated <- c(BCG.Unvaccinated, BCG.vaccinations[4:6]* 4/3*0.11)
+names(BCG.Unvaccinated) <- c('1967-71','1972-1976','1977-1981','1982-86','1987-91','1992-1996')
+
+################# No. of Tuberculin postive people: set to be average of collected data at 8%
+BCG.Tuberculin <- c(0.336*0.95, 0.291 * 0.943, 0.241 * 0.936)
+BCG.Tuberculin <- BCG.Tuberculin*10^6
+BCG.Tuberculin <- c(BCG.Tuberculin, BCG.vaccinations[4:6]* 4/3*0.08)
+names(BCG.Tuberculin) <- c('1967-71','1972-1976','1977-1981','1982-86','1987-91','1992-1996')
+
+################# No. of Tuberculin postive people: set to be average of collected data at 6 %
+BCG.Inelgible <- c(0.248*0.95, 0.201 * 0.943, 0.155 * 0.936)
+BCG.Inelgible <- BCG.Inelgible*10^6
[…]

### 2019-04-10 data-raw/sutherland_gen_time.R

@@ -1,0 +1,6 @@
+## Assumption from sutherland et al.
+sutherland_gen_time <- 2
+
+#' Add to package and keep in data-raw
+devtools::use_data(sutherland_gen_time, overwrite = TRUE)
+readr::write_csv(as.data.frame(sutherland_gen_time), "sutherland_gen_time.csv")

### 2019-04-10 data-raw/sutherland_incidence_rates.R

@@ -1,0 +1,34 @@
+Cohort <- c('1967-71','1972-1976','1977-1981')
+Cohort.Ranges <- c('15-19 years', '20-24 years', '25-29 years')
+
+### Assumed Incidence rates from sutherland et al.
+#################### TB rates in unvaccinated people per 100,000
+Unvaccinated <- t(matrix(c(22.88, 20.06, 11.58, 10.81, 11.42, 0, 7.59, 0, 0),3,3))
+rownames(Unvaccinated) <- Cohort
+colnames(Unvaccinated) <- Cohort.Ranges
+
+#################### TB rates in vaccinated people per 100,000
+Vaccinated <- t(matrix(c(2.57,4.446,3.83,2.70,3.12,0,1.53,0,0),3,3))
+rownames(Vaccinated) <-  Cohort
+colnames(Vaccinated) <-  Cohort.Ranges
+
+#################### Notifcations prevented per 100,000
+Notifications.Prevented <- Unvaccinated-Vaccinated
+
+################# TB rates in Tuberculin positive people
+Tuberculin <- t(matrix(c(19.98, 7.21, 3.83, 14.34, 3.02, 0, 4.5, 0, 0),3,3)) # Real Notifcations rate from previous papers
+rownames(Tuberculin) <- Cohort
+colnames(Tuberculin) <-Cohort.Ranges
+
+################# TB rates in Inelgible people - assumed to be equal to those in the unvaccinated as not stated in
+## published work
+
+Inelgible  <- Unvaccinated
+
+sutherland_incidence_rates <- list(Unvaccinated, Vaccinated, Notifications.Prevented, Tuberculin, Inelgible)
+names(sutherland_incidence_rates) <- c("Unvaccinated", "Vaccinated", "Notifications.Prevented", "Tuberculin", "Inelgible")
+
+
+#' Add to package and keep in data-raw
+devtools::use_data(sutherland_incidence_rates, overwrite = TRUE)
+save(sutherland_incidence_rates, file = "sutherland_incidence_rates.rda")

### 2019-04-10 data-raw/sutherland_results.R

@@ -1,0 +1,13 @@
+## Sutherland results
+## Total notifications from ending the scheme
+sutherland_results <- matrix(c(288, 288, 288, 226, 165, 165, 208,
+                                           130,  90, 181, 128,  77, 128, 115,
+                                           80,  80,  80,  72), 3, 6)
+
+colnames(sutherland_results) <- c('1988', '1993', '1998', '2003', '2008', '2013')
+rownames(sutherland_results) <- c('1986','1991','1996')
+
+
+#' Add to package and keep in data-raw
+devtools::use_data(sutherland_results, overwrite = TRUE)
+readr::write_csv(as.data.frame(sutherland_results), "sutherland_results.csv")

### 2019-04-10 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(AssessBCGPolicyChange)
+
+test_check("AssessBCGPolicyChange")

### 2019-04-10 README.Rmd
bumped doi link

@@
 [![Paper](https://img.shields.io/badge/Paper-10.1101/567511-lightgreen.svg)](https://doi.org/10.1101/567511)
-[![DOI](https://zenodo.org/badge/173767331.svg)](https://zenodo.org/badge/latestdoi/173767331)
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2635687.svg)](https://doi.org/10.5281/zenodo.2635687)
 

### 2019-04-29 peer-review/cover-letter.Rmd
Updated cover-letter based on E feedback

@@
 
-Please find attached our manuscript, entitled "Reassessing the Evidence for Universal School-age Bacillus Calmette Guerin (BCG) Vaccination in England and Wales".
+Please find attached our manuscript, entitled “Reassessing the Evidence for Universal School-age Bacillus Calmette Guerin (BCG) Vaccination in England and Wales” for consideration for publication in the journal of Epidemiology and Community Health.
 
-We feel our manuscript will be of interest to you and the journal's readership because Bacillus Calmette-Guérin (BCG) is one of the most widely-used vaccines worldwide and our study explores the impact of changing BCG vaccination policy.
-We found that previous estimates of the impact of ending the BCG schools' scheme may have been spuriously precise as parameter uncertainty and measurement error were not fully included.
-Using surveillance data, we updated these estimates and found that the impact of ending the BCG schools' scheme was greater than previously estimated.
-These results highlight the importance of including parameter and measurement error when forecasting the impact of changes in vaccination policy.
-They also highlight the benefits of reproducibility as our study built on work previously published in this journal.
-This study benefited from access to previously published work that was used by policy makers.
-We combined this with more modern statistical methods to fully capture uncertainty and used recent data to increase the accuracy of the model estimates.
+In this paper, we examine a previous analysis of the impact of stopping universal vaccination in England published in the journal of Epidemiology and Community Health in 1989 in light of changing TB epidemiology.
+Using recent surveillance data, together with modern statistical methods to capture uncertainty, we found that the impact of ending the BCG schools’ scheme was greater than previously estimated.
+Previous estimates were spuriously precise as parameter uncertainty and measurement error were not fully included.
+Our work highlights the benefits of reproducibility as our study built on work previously published in this journal.
 
+We believe our manuscript will be of interest to you and the journal’s readership because Bacillus Calmette-Guérin (BCG) is one of the most widely-used vaccines worldwide and our study explores the impact of changing BCG vaccination policy.
 We do not have any issues relating to the journal policies.

### 2019-05-01 peer-review/cover-letter.Rmd
added wording changs to draft
added updates for phrasing prior to ubmission

@@
 
-In this paper, we examine a previous analysis of the impact of stopping universal vaccination in England published in the journal of Epidemiology and Community Health in 1989 in light of changing TB epidemiology.
+In this paper, we examine a previous analysis of the impact of stopping universal vaccination in England published in the journal of Epidemiology and Community Health in 1989 in light of changing Tuberculosis epidemiology.
 Using recent surveillance data, together with modern statistical methods to capture uncertainty, we found that the impact of ending the BCG schools’ scheme was greater than previously estimated.
 Previous estimates were spuriously precise as parameter uncertainty and measurement error were not fully included.
-Our work highlights the benefits of reproducibility as our study built on work previously published in this journal.
+Our work highlights the benefits of reproducibility as our study was built on work previously published in this journal.
 
 We believe our manuscript will be of interest to you and the journal’s readership because Bacillus Calmette-Guérin (BCG) is one of the most widely-used vaccines worldwide and our study explores the impact of changing BCG vaccination policy.
+
 We do not have any issues relating to the journal policies.

### 2019-05-01 README.Rmd
readme updates

@@
 
-In 2005, England and Wales switched from universal BCG vaccination for school-age children to targeted vaccination of neonates.
-We assessed some of the quantitative evidence that informed this policy change.
+In 2005, England and Wales switched from universal BCG vaccination against tuberculosis (TB) disease for school-age children to targeted vaccination of neonates.
+We assessed the quantitative evidence that informed this policy change.
 
@@
 
-We recreated a previous approach for estimating the impact of ending the BCG schools scheme in England and Wales, updating the model with parameter uncertainty.
-We investigated scenarios considered by the Joint Committee on Vaccination and Immunisation, and explored new approaches using notification data.
+We recreated a previous approach for estimating the impact of ending the BCG schools’ scheme in England and Wales, updating the model with parameter uncertainty.
+We investigated scenarios considered by the UK’s Joint Committee on Vaccination and Immunisation, and explored new approaches using notification data.
 We estimated the number of vaccines needed to prevent a single notification, and the average annual additional notifications caused by ending the BCG schools’ scheme.
@@
 
-We found a 1.9% annual decrease in Tuberculosis incidence rates best matched our estimates from notifications.
-We estimate that 1600 (95% Interquartile range (IQR) 1300 to 2100) vaccines would have been required to prevent a single notification in 2004.
-If the scheme had ended in 2001, 302 (95% IQR 238 to 369) additional annual notifications would have occurred compared to if the scheme had continued.
-If the scheme ended in 2016, 120 (95% IQR 88 to 155) additional annual notifications would have occurred.
+We found a 1.9% annual decrease in TB incidence rates best matched notification data.
+We estimate that 1600 (2.5-97.5% Quantiles (Q): 1300 - 2100) vaccines would have been required to prevent a single notification in 2004.
+If the scheme had ended in 2001, 302 (2.5-97.5% Q: 238 - 369) additional annual notifications would have occurred compared to if the scheme had continued.
+If the scheme ended in 2016, 120 (2.5-97.5% Q: 88 - 155) additional annual notifications would have occurred.
 
@@
 
-Our estimates of the impact of ending the BCG schools scheme were highly sensitive to the annual decrease in incidence rates.
-The impact of ending the BCG schools scheme was found to be greater than previously thought when parameter values were updated and notifications were used.
+Our estimates of the impact of ending the BCG schools’ scheme were highly sensitive to the annual decrease in incidence rates.
+The impact of ending the BCG schools’ scheme was found to be greater than previously thought when parameter values were updated and notification data were used.
 Our results highlight the importance of including uncertainty when forecasting the impact of changes in vaccination policy.

### 2019-05-07 README.Rmd
updated DOI for preprint

@@
 [![Documentation](https://img.shields.io/badge/Documentation-click here!-lightgrey.svg?style=flat)](https://www.samabbott.co.uk/AssessBCGPolicyChange)
-[![Paper](https://img.shields.io/badge/Paper-10.1101/567511-lightgreen.svg)](https://doi.org/10.1101/567511)
+[![Paper](https://img.shields.io/badge/Paper-10.1101/624916-lightgreen.svg)](https://doi.org/10.1101/624916)
 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2635687.svg)](https://doi.org/10.5281/zenodo.2635687)

### 2019-05-07 README.Rmd
updated readme

@@
 
-Alternatively the analysis environment can be accessed via [binder](https://mybinder.org/v2/gh/seabbs/DirectEffBCGPolicyChange/master?urlpath=rstudio).
-
-# Docker
-
-This packge was developed in a docker container based on the tidyverse docker image. To run the docker image run:
-
-```{bash, eval = FALSE}
-docker run -d -p 8787:8787 --name assessbcgpolicychange -e USER=seabbs -e PASSWORD=seabbs seabbs/assessbcgpolicychange
-```
-
-The rstudio client can be found on port :8787 at your local machines ip.
-The default username:password is seabbs:seabbs, set the user with -e USER=username, and the password with - e PASSWORD=newpasswordhere.
-The default is to save the analysis files into the user directory.
+Alternatively the analysis environment can be accessed via [binder](https://mybinder.org/v2/gh/seabbs/AssessBCGPolicyChange/master?urlpath=rstudio).

### 2019-05-09 peer-review/cover-letter.Rmd
changes for BMJ open

@@
 
-Please find attached our manuscript, entitled “Reassessing the Evidence for Universal School-age Bacillus Calmette Guerin (BCG) Vaccination in England and Wales” for consideration for publication in the journal of Epidemiology and Community Health.
+Please find attached our manuscript, entitled “Reassessing the Evidence for Universal School-age Bacillus Calmette Guerin (BCG) Vaccination in England and Wales” for consideration for publication in BMJ Open.
 
@@
 Previous estimates were spuriously precise as parameter uncertainty and measurement error were not fully included.
-Our work highlights the benefits of reproducibility as our study was built on work previously published in this journal.
+Our work highlights the benefits of reproducibility as our study was built on a previously published model.
 

### 2019-05-10 README.Rmd
further updates for BMJ open

@@
 
-## Background
+## Objectives
 
 In 2005, England and Wales switched from universal BCG vaccination against tuberculosis (TB) disease for school-age children to targeted vaccination of neonates.
-We assessed the quantitative evidence that informed this policy change.
+We aimed to assess the quantitative evidence that informed this policy change.
 
-## Methods
+## Design
 
-We recreated a previous approach for estimating the impact of ending the BCG schools’ scheme in England and Wales, updating the model with parameter uncertainty.
+We recreated a previous approach for estimating the impact of ending the BCG schools’ scheme, updating the model with parameter uncertainty.
 We investigated scenarios considered by the UK’s Joint Committee on Vaccination and Immunisation, and explored new approaches using notification data.
-We estimated the number of vaccines needed to prevent a single notification, and the average annual additional notifications caused by ending the BCG schools’ scheme.
+
+## Setting
+
+England and Wales.
+
+## Outcome measures
+
+The number of vaccines needed to prevent a single notification, and the average annual additional notifications caused by ending the BCG schools’ scheme.
 
@@
 We found a 1.9% annual decrease in TB incidence rates best matched notification data.
-We estimate that 1600 (2.5-97.5% Quantiles (Q): 1300 - 2100) vaccines would have been required to prevent a single notification in 2004.
-If the scheme had ended in 2001, 302 (2.5-97.5% Q: 238 - 369) additional annual notifications would have occurred compared to if the scheme had continued.
-If the scheme ended in 2016, 120 (2.5-97.5% Q: 88 - 155) additional annual notifications would have occurred.
+We estimate that 1600 (2.5-97.5% Quantiles (Q): 1300-2100) vaccines would have been required to prevent a single notification in 2004.
+If the scheme had ended in 2001, 302 (2.5-97.5% Q: 238-369) additional annual notifications would have occurred compared to if the scheme had continued.
+If the scheme ended in 2016, 120 (2.5-97.5% Q: 88-155) additional annual notifications would have occurred.
 
@@
 Our estimates of the impact of ending the BCG schools’ scheme were highly sensitive to the annual decrease in incidence rates.
-The impact of ending the BCG schools’ scheme was found to be greater than previously thought when parameter values were updated and notification data were used.
+The impact of ending the BCG schools’ scheme was found to be greater than previously thought when parameter values were updated, and notification data were used.
 Our results highlight the importance of including uncertainty when forecasting the impact of changes in vaccination policy.

### 2019-09-03 peer-review/review-response.Rmd
outlined reviewer responses

---
title: "Reviewer Response"
output:
  word_document: default
  html_document:
    df_print: paged
---

Dear Editor,

Thank you for sending review reports for our manuscript. The reviewer’s comments were useful and have improved the paper.

We enclose a revised version of the manuscript (clean and with tracked changes), which addresses the reviewer’s comments and suggestions. The reviewer’s comments are shown below with our response to each comment in italics.

Yours sincerely,

Sam Abbott

## Authors' response to Editorial Requests

- Please include the study design in the title.

**What do we think the study design is here?**

- In the patient and public involvement statement, please clarify that patients and the public were not involved in the design or planning of the study.

*We have revised the patient and public involvement statement to the following,*

*"We did not involve patients or the public in design or planning of this study."*

- Please complete and include a TRIPOD checklist, ensuring that all points are included and state the page numbers where each item can be found. The checklist can be downloaded from here: http://www.equator-network.org/reporting-guidelines/tripod-statement/

**I'm not sure this checklist is really appropriate for this paper. It is primarly designed for stastical models used for prediction.**

## Authors' response to Reviewer #1:

This study revisited the model used for driving the change of BCG vaccination policy in 2005 in the UK. Assumptions on the declining trend of TB burden were evaluated and found to be an important source of uncertainty in estimating the impacts of BCG vaccination. This work will contribute to the understanding of appropriate evidence needed for planning BCG vaccination strategies. My comments for the manuscript are listed below.

*Thank you for your helpful comments on this paper.*

### Introduction

1.	Page 4, Line 21: BCG efficacy has been known with great heterogeneity and it affects the assessment of vaccine impact on TB burden. It is therefore commonly considered in decision-making for BCG vaccination policy. I suggest that the authors include the uncertainty of BCG efficacy in this analysis and compared the relative importance to the assumption of how TB burden declines over time.

*The reviewer is correct in identfying this as an issue. We have included some of this uncertainty by re-estimating incidence rates for both those BCG vaccinated and unvaccinated (this is uncertainty is then passed through to the efficacy estimates). This approach makes use of the available data on the effiacy of the BCG vaccine in England and Wales. This does mean that the findings from this model reflect the impact in these countries only and cannot be directly generalised.*

2.	Page 4, Line 36: The abbreviation should be ‘JCVI’ rather than ‘JVCI’. Note that the same typo appears in other paragraphs of the manuscript as well as the figure captions.

*Thank you. We have corrected this throughout*

### Methods

3.	Page 5, Line 6-7: Is the ‘transmission chain model’ included in the Sutherland et al.’s model or developed by the authors? Please state it here.

*We have extensively reworded and extended this section (please see the marked up version). We have added the following sentence for clarity,*

*"The TB transmission model is defined as follows (reproduced from Sutherland et al.)"*

4.	Page 5, steps for defining the transmission model: Presenting equations of the model is certainly helpful for understanding the estimation methods. In addition to the descriptions of equation components, I encourage the authors to explain more about the rationales behind. For example, how to interpret ‘initial generation size (x)’? What is the purpose of relating step 1 to step 2? This kind of additional information will allow readers from a wide range of background to understand the model easily.

*Whilst this information was present in the original Sutherland et al. paper we agree that more detailed was needed here. We have therefore added greater explanatory detail for each step (please see the marked up paper).*

5.	Page 5, last paragraph: Is the ‘distribution of secondary cases (N)’ estimated by the three age groups (of 15-19, 20-24, and 25-29 year old)? The power term ‘5-z’ of the decay percentage (d) needs additional explanations in the text.

*We agree. We have reformulated these equations for greater clarity (please see the marked up paper).*

6.	Page 5, Line 54-55: Please revise the sentence fragment.

*Revised to:*

*"We validated the fitted model by comparing the results with those from the original implementation using the mean absolute percentage error, normalised by the original estimate, as the performance metric."*

7.	Page 6, Line 10-11: Is the serial interval defined as between linked ‘notifications’, instead of ‘infections’?

*Thank you - corrected.*

8.	Page 6, Line 28-32: This part is too long to follow and with sentence fragments. Please re-organise it.

*Reworded to:*

*"Data collection for the ETS began in 2000 and prior to this notification data was only available in years with notifications surveys (1973, 1978, and 1983). We therefore estimated incidence rates between 1984 and 1999, and for the years between notifications surveys (1974-1977 and 1979-1982), using LEOSS regression fitted incidence rates published in [11], and the estimated incidence rates from 2000 on-wards. LOESS (locally estimated scatterplot smoothing) is a local regression method.[@FoxSnow:2019] It combines multiple regression models in a k-nearest neighbours meta-model. This approach allows nonlinear trends to be fitted to using a series of linear models. For years prior to 1973 the annual decreases were assumed to be the mean of the previous 3 years of data. For both proxy measures the annual decreases in incidence rates post 2016 were assumed to be the average of the estimates in 2013, 2014, and 2015."*

### Results

9.	Page 8, Line 8-10: ‘The large degree of uncertainty’ was related to the different assumptions of the annual TB decrease. However, here it sounds like the uncertainty is in the range of a particular scenario estimate.

*The range of uncertainty here is within the scenario. This is due to the incorperation of uncertainty detailed in the Updating model parameter estimates section of the methods.*

### Discussion

10.	Page 8, Line 40-41: ‘A scenario with a 1.9% annual decrease in incidence rates was most comparable to our results based on notifications.’ As the medium annual decrease estimated using notifications was 3.13% (Page 7, Line 20), why the scenario of a 3.9% annual decrease did not give the most comparable results?

*Unfortunately, the summary statistics do not capture changes in the annual decrease in notifications over time. Notifications first increased and then decreased and it is these trends that led to the overall 1.9% decrease being a better fit. To help clarify this we have moved a figure from the SI into the main body of the paper.*

11.	Page 9, Line 37-38: Do you consider to report the primary impacts of ending the BCG programme and the additional cases due to transmission separately? This will show the importance of including transmission structure in assessing the BCG vaccination policy.

*We agree that including transmission structure is important when assessing the impact of infectious disease programmes. We did not consider seperating these estimates as we felt it was well established in the literature that transmission structure should always be included. Whilst adding to this body of evidence would be worthwhile we feel that adding additional outcome measures may further complicate the results section of this paper.*

## Authors' response to Reviewer #2:

### OVERVIEW

This paper revisits an analysis that was used to support the decision to discontinue the UK policy of BCG vaccination in schools. It considers updated evidence on parameters as well as incorporating uncertainty in model inputs. Re-evaluating modelling around decisions and considering the impact of uncertainty are both under-explored and worthy topics.

At the moment, I think this article needs major revisions before publication as it is extremely difficult to follow exactly what has been done.

*Thank you for your comments. Based on your them we have greatly extended our discussion of the methods used and added further clarification where possible. As this paper reproduces the model published by Sutherland et al. some elements cannot be updated without making the findings difficult to compare. Please see our detailed responses below.*

### MAJOR ISSUES

The idea of the approach is that information about the rate of decline of TB incidence over time combined with knowledge of the generation time should inform on the net reproduction number, and hence the scope for indirect benefits from preventing cases. However, I really can't follow the model description on page 5. Part 1: T (and in part 2) is described as the sum over all generations of resulting cases, but R0^z (presumably R rather than R0) and (1-d)^z seem very much like the expected size of the z-th generation of resulting cases? Part 2: x is described as an initial size, but it seems in fact to be a net reproduction number from the way it is used in a geometric progression (resulting cases should scale linearly with initial size under branching process type models). If x=R and eqn 1 is true, things seem over-determined: one can solve R(z) without need for data on d?? Part 3 makes sense if x is R. Then I also don't understand how this relates to the 3 equations at the bottom of the page. I also don't remember the eligibility for BCG vaccination being introduced (line 37), and am not clear what it means to 'reproduce the distribution of cases due to transmission over time'. I think all this needs much more careful explanation.

*We agree that this needed further explanation. We have greatly extended the explanation using information from the SI and added further clarification where possible.*

I also found it harder than it should have been to understand exactly what was different between this analysis and the original analysis, and which differences mattered most (which is important as it is the main aim of the work as I understand it). I wonder if a table would be useful detailing the differences: any implementation/modelling differences; differences in point estimates of d (both from methods and use of more recent data?); differences in parameters from literature; inclusion of uncertainty in each. I think it's important to differentiate between differences that could have been included in the analysis at the time (eg uncertainty) vs things that have emerged since (eg updated evidence on serial intervals). I also think it's important to be clear whether methods, central estimates, or uncertainty matter most from the point of view of decision-relevant changes in the evidence generated.

*To answer your previous comment we introduction a subsection of the model explanation detailing our additions to the model. The Updating model parameter estimates section of the methods contains a discussion of differences between the original and updated analysis. We have reworded this for clarity (please see the marked up paper). As you say, parameter uncertainty for incidence rates could have been included at the time but was not, whilst the updated value for the serial interval has only recently become available. This is now explicitly stated in the rewored methods paragraph.*

*We agree that it is important to be clear which findings are relevant to policy makers and in what context. We have rewored the 4th discussion paragraph as follows in order to provide more context.*

*"This study indicates that some of the evidence used to justify the 2005 change in BCG vaccination policy may have underestimated the impact of ending the scheme. It highlights the importance of including both parameter and measurement error, as excluding these sources of variation may lead to spuriously precise results. This is problematic for policy makers as the worst case scenario often needs to be considered. In addition, my exploration of the assumptions used to estimate the annual change in TB incidence rates in the ethnic White UK born illustrates the structural impact of assuming an annual decrease in TB incidence rates. More realistic estimates of the annual decrease in incidence rates resulted in a greatly increased impact of ending the BCG schools scheme. Policy makers should consider these updated estimates when assessing the role of BCG vaccination in those at school-age. However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiolgical impact of vaccination only."*

I found the abstract confusing and not self-contained. "assess the quantitative evidence..." is rather vague and broad compared with my perception of the paper's aim. I think the fact a model was used to predict impact needs introducing, as it enters rather implicitly. It would be good to say in what ways the analysis was updated (see previous comment). "best matched" is vague. "ended in 2001" - need to have explained when it did end. Why is 2004 important? The results presented here are all about different end-dates for the vaccination policy rather than comparisons with the original analysis's predictions. "In 2016...notifications would have occurred." - under what policy change? In the conclusions: it hasn't be explained that decline in incidence rates is relevant to the modelling, it hasn't been made clear that parameters have been udated and a comparison with previous results has not been presented. In other words, the conclusion introduces new things, which is a shock, and is not supported by the rest of the abstract.

**Abstract updates still a work in progress*.*

### MINOR ISSUES

- " BCG schools' scheme" in many places. Is this what it was officially called? I didn't understand why it had an apostrophe.

*Thank you - corrected throughtout.*

- In a couple of places (ie strengths/limitations bullets & Discussion) it is stated this is not "the most accurate method" for assessing the impact of these changes. This does beg questions about what other methods would have been better, and in what respects (eg how is "accuracy" defined?). This is worth discussing. I think it also means that the rationale for the work is about assessing the imapct of updated/modified analysis for inputs within an existing framework, rather than trying to provide a better answer to the same question (otherwise, why use a suboptimal method?)

*We agree that this could have been better phrased. We have rephrased as follows,*

*"The simulation approach used here is not the most accurate method for estimating the impact of ending the BCG schools’ scheme as it relies on numerous assumptions based on the available knowledge in 1987. The strength of this work is that the estimates were based on the framework used to inform policy making. This allowed the strength of model used in the decision-making process to be assessed once parameter uncertainty had been incorperated. This would not have been possible if the impact had been assessed using only the observed data. It also allowed estimates based on updated data to be compared to historic estimates within the same framework. If a different framework had been used then these estimates would not have been comparable."*

*We discuss alternative approaches for assessing the impact in the final paragraph of the discussion but you are correct in saying that the rationale of this work as to assess the impact of update an existing framework. We have clarified this by adding the following sentence to the introduction,*

*"Re-evaluating this work allows for the strength of the evidence used in decision making to be assessed and aims to highlight any issues with the approach used."*

- page 4 line 36: missing apostrophe in "papers"

*Corrected.*

- last 2 sentences of intro are Discussion

*Removed.*

- "TB notification surveys" - could you explain what these are please?

*Added detail from the SI into the main body of the methods. Please see the marked up paper.*

- ref 17: I thought there had been a systematic review of serial intervals for TB (Ma et al, Epidemiology & Infection). Justify why using this study in particular please. Choice of evidence an important part of paper since arguing to update & compare.

*Thanks for pointing this out. Whilst there has been a systematic review the authors identifed only two papers with estimates (the one used here and one other older paper). Both papers made use of similar datasets in the Netherlands. The cited paper is 10 years newer, has an improved sample size, and improved methods.*

- pg 6 paragraph on methods to quantify decline: again, I found this confusing and it is a key part of the analysis. Also, what is "local regression" please?

*We have reworded this paragraph (and made changes suggested by reviewer 1). Please see the marked up paper. We have also included a graph in the results from the SI that shows  the changes over time in the annual decline depending on the data/assumptions used.*

*We have added the following clarification for local regression + reference,*

*"LOESS (locally estimated scatterplot smoothing) is a local regression method.[@FoxSnow:2019] It combines multiple regression models in a k-nearest neighbours meta-model. This approach allows nonlinear trends to be fitted to using a series of linear models."*

- pg 7 paragraph on incidence rate changes: many medians presented do not lie between the corresponding 2.5% and 97.5% quantiles, which must be typos. Also there seems to be some confusion about whether to present decreases or negative increases.

*Unfortunately, our choice of "-" to seperate estimates the 2.5% and 97.5% quantile made understanding this paragraph difficult. We have moved to using "," as this morely identifies when estimates are negative. Thank you for picking this up.*

- NNT vaccination results. I'm surprised by the factor of 10 change in the NNV going from 9% decline to 2% decline. Could this be double checked please? Sanity check: if TB incidence is of the order 10/100,000 and the vaccine better than 50% effective, I imagine NNV is ought to be less than 20,000 considering only individual-benefit (the magnitude calculated for 9% decline); to achieve the stated effect must involve many generations of transmission (perhaps exceeding the time-frame considered?). However, it would be useful to clearly ellaborate this: perhaps a table of these results, with proportion of benefit from individual vs transmission and estimates of R for each decline considered; so the way the numbers hang together can be more clearly understood.

*We agree that these estimates (from the original Sutherland et al. paper) are very unrealistic. Sutherland et al. assumed a year-on-year decline of 9% with an estimated incidence of around 23 per 100,000 in 1983 for the unvaccinated population. In 2013 this would give an incidence rate of around 2.6. These very small incidence rates mean that the difference between incidence rates in unvaccinated and vaccinated cases is also very small leading to the spuriously high numbers need to vaccinate. In comparison a 2% decline would give an inidence rate of 12.5. Our additions to the methods section add extra detail on how these estimates are arrived at.*

- I think discounting needs to be discussed at some point. This is almost always included in decision analyses.

*We agree that discounting should be included for economic evaluation. However we have not seen it done widely when the study is focussing purely on the epidemiological impact of vaccination. We have added the following sentences to the discussion,*

*"Policy makers should consider these updated estimates when assessing the role of BCG vaccination in those at school-age. However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiolgical impact of vaccination only."*

- Figures 1 & 2: "Notifications" & "Incidence Rates" here are methods to derive a rate of annual decline. I think it would make more sense to label these with the derived rate of decline, order the y axis by these, and perhaps use color to label the method used to drive the rates of decline.

*The derived rates are variable by year and hence cannot be used as a label. We have included a figure from the SI in order to clarify this.*

- I wonder whether it is possible to compare the predicted effect on TB notifications made at the time of the policy change to actual observations and comment on this? If not, why?

*The model presented here is based on a series of assumptions from the 1980s that have not been updated (such as the population size) and does not account for notifications in the non-UK born and non-White populations which are likely to have a significant impact on incidence rates in the White UK born. For these reasons we have only compared between model estimates in this study as the primary aim was to assess the strength of the model used in decision making. As mentioned above we have clarified the aim of the study in the introduction and abstract.*

## Authors' response to Reviewer #3:

This manuscript reports the results of a mathematical modeling analysis that updates a previous analysis done in 1987 to project the potential effect of ending universal school-age vaccination in the UK. It uses the same model framework but incorporates uncertainty estimates (which the previous model did not) and explores the effect of changing one of the original model assumptions about rate of decline in TB incidence to what has actually been observed in the intervening decades.

*Thank you for your comments.*

I think that the authors have executed their stated goal well, but I am having a hard time seeing this as an independently publishable piece of work. Making minor tweaks to the model parameters and adding uncertainty estimates did not produce fundamentally new conclusions. The more important change in the assumption of rate of decline in TB incidence did change the results, but the conclusions still do not seem independently publishable; the conclusion seems to be simply that the projected impact depends very much on what you assume is happening with the epidemic. I wonder if maybe I am not understanding the results completely? It seems like since 15 years have passed since the policy change, the more useful analysis would be to try to assess what the impact has been in those 15 years rather than repeat the projection exercise that led to the policy change with varying assumptions about declines in TB incidence when in fact it should be knowable what has happened in those 15 years. Maybe this is what the authors were trying to do, but I find myself getting to the end of the article with the unanswered question about whether the policy change was the right decision or not.

*We agree that this study does not fully answer whether or not the policy change was the correct decission and agree that analysising recent data may be a more valuable approach for this (https://doi.org/10.1101/567511). The aim of this study was instead to assess the strength of the evidence used for decision making at the time of the policy change - we have updated the abstract and introduction to clarify this. As mentioned by the second reviewer models are rarely evaluated once the policy decision has been made. In this study we show that the model used for deicision making was spuriously precise and made several assumptions that led to what we now know were entirely unrealistic estimates. These findings are important for decision makers to understand when evaluating quantative evidence for future policy decisions.*

I think I found this paper difficult to understand because it referred so heavily to the older piece of work. I understand that this paper is updating the previous results, but it is my opinion that every scientific paper has to stand alone. Without a conventional diagram of the model structure, list of parameters with sources, and definitions of key terms, I had a hard time piecing together what was done, and I am still not sure that I fully understand. Some of this may reflect my own lack of expertise in building these kinds of models, so I realize that to a reader more in the authors’ field, this may not have been as much of an issue. However, in the interest in making this work more accessible to the general reader (and reviewer), some specific key terms and procedures that I did not find well described are listed below:

*We agree that without the context of the SI and the Sutherland et al. paper the methods section was difficult to follow.  We have greatly expanded on the explanation of the model given in the methods incorperating details from the SI and improved the discussion of the Sutherland et al. model.*

- Introduction paragraph 2: Does “BCG schools’ scheme” mean universal BCG for school children? What age were children vaccinated? What was coverage assumed to be – that is, were any children assumed to be unvaccinated during the old policy? After the policy change, did anyone get vaccinated (e.g. high-risk children)?

*We have expanded on this in the introduction as follows, *

*"In England and Wales, universal school-aged (at 13-14 years old) vaccination was introduced after a MRC trial in the 1950s estimated BCG’s effectiveness at 78% in the ethnic White UK born population.[24] The policy remained in place until 2005, when England and Wales changed to targeted vaccination of high risk neonates. The 2005 change in BCG vaccination policy was motivated by evidence of decreased transmission of TB, an increasing proportion of TB cases occurring in the non-UK born,[1] and modelling evidence that suggested stopping the BCG schools scheme would have minimal long term effects on incidence rates.[39]"*

- Methods, paragraph 1: Who are the vaccinated and unvaccinated populations? Is vaccinated = hypothetical situation of BCG school vaccinations having continued, and unvaccinated = situation in which policy changed to stop vaccinations? Or in the two situations are different proportions of children assumed to be vaccinated and unvaccinated?

*We have incorperated details from the SI in order to expand on this - please see the marked up paper.*

- Methods, paragraph about assumptions: “Firstly, as incidence rates for those ineligible for the BCG schools scheme are not published…” Who was ineligible if it was a universal program as stated in the introduction? The actual policies need to be better described.

*See above.*

- It seems that somewhere in the model the estimated protective effect of BCG should appear, unless I am not understanding the model structure at all (which is possible).

*See above.*

- Methods, section on updating model parameter estimates: Were the data used for estimating changes in incidence obtained only for the white non-foreign-born population or for the entire English population? Also, it is unclear what these two data sources are (i.e. what they capture) and how they differ. The limitations suggest that maybe the notification data were not race/origin stratified and something was done to account for this, but the methods do not describe this.

*We apologise for the confusion. Data on incidence rates in the White UK born was not available. Therefore, we used two proxy measures notifications in england and Wales and incidence rates in the UK born in England only (data for Wales was not available). Based on your comments (and those of the other reviewers) we have reworded this paragraph.*

- Results: the statement is made that “We found that updating parameter values, and incorporating uncertainty, did not alter the number of vaccines required to prevent a single notification…” But changing the assumed decline in incidence rate does have a major impact. Isn’t this one of the parameters of the model? Or am I not understanding how the model works?

*We apologise for the confusion - hopefully our updates to the methods section has helped clarify the model structure. We considered the annual decrease in a scenario analysis but all other parameters were otherwise updated.*

### 2019-09-04 peer-review/review-response.Rmd
initial reviewer response

@@
 This approach makes use of the available data on the effiacy of the BCG vaccine in England and Wales.
+Including additional uncertainty observed globally would be inappropriate here as there is little evidence for this amongst the White UK born population.
 This does mean that the findings from this model reflect the impact in these countries only and cannot be directly generalised.*
@@
 
-**Abstract updates still a work in progress*.*
+**Abstract updates still a work in progress.**
 
@@
 
-*We discuss alternative approaches for assessing the impact in the final paragraph of the discussion but you are correct in saying that the rationale of this work as to assess the impact of update an existing framework.
+*We discuss alternative approaches for assessing the impact in the final paragraph of the discussion but you are correct in saying that the rationale of this work was to assess the impact of updateingree an existing framework.
 We have clarified this by adding the following sentence to the introduction,*
 
-*"Re-evaluating this work allows for the strength of the evidence used in decision making to be assessed and aims to highlight any issues with the approach used."*
+*"Re-evaluating this work allows for the strength of the evidence used in decision making to be assessed and may highlight any issues with the approach used."*
 
@@
 *Unfortunately, our choice of "-" to seperate estimates the 2.5% and 97.5% quantile made understanding this paragraph difficult.
-We have moved to using "," as this morely identifies when estimates are negative.
+We have moved to using "," as this more clearly identifies when estimates are negative.
 Thank you for picking this up.*
@@
 
-*We agree that this study does not fully answer whether or not the policy change was the correct decission and agree that analysising recent data may be a more valuable approach for this (https://doi.org/10.1101/567511).
+*We agree that this study does not fully answer whether or not the policy change was the correct decision and agree that analysising recent data may be a more valuable approach for this (see here for a preprint by 3 of this studies authors doing this: https://doi.org/10.1101/567511).
 The aim of this study was instead to assess the strength of the evidence used for decision making at the time of the policy change - we have updated the abstract and introduction to clarify this.

### 2019-09-05 peer-review/review-response.Rmd
further reviewer changes

@@
 *Whilst this information was present in the original Sutherland et al.
-paper we agree that more detailed was needed here.
+paper we agree that more detail was needed here.
 We have therefore added greater explanatory detail for each step (please see the marked up paper).*
@@
 *"Data collection for the ETS began in 2000 and prior to this notification data was only available in years with notifications surveys (1973, 1978, and 1983).
-We therefore estimated incidence rates between 1984 and 1999, and for the years between notifications surveys (1974-1977 and 1979-1982), using LEOSS regression fitted incidence rates published in [11], and the estimated incidence rates from 2000 on-wards.
-LOESS (locally estimated scatterplot smoothing) is a local regression method.[@FoxSnow:2019] It combines multiple regression models in a k-nearest neighbours meta-model.
-This approach allows nonlinear trends to be fitted to using a series of linear models.
+We therefore estimated incidence rates between 1984 and 1999, and for the years between notifications surveys (1974-1977 and 1979-1982), using locally estimated scatterplot smoothing (LEOSS) regression fitted to incidence rates published in [11] and the estimated incidence rates from 2000 on-wards.
+LOESS is a local regression method that combines multiple regression models in a k-nearest neighbours meta-model.[@FoxSnow:2019] This approach allows nonlinear trends to be fitted to using a series of linear models.
 For years prior to 1973 the annual decreases were assumed to be the mean of the previous 3 years of data.
@@
 
-*Thank you for your comments.
-Based on your them we have greatly extended our discussion of the methods used and added further clarification where possible.
-As this paper reproduces the model published by Sutherland et al.
-some elements cannot be updated without making the findings difficult to compare.
-Please see our detailed responses below.*
+*Thank you for your comments. Based on them we have greatly extended our discussion of the methods used and added further clarification where possible. Please see our detailed responses below.*
 
@@
 
-*To answer your previous comment we introduction a subsection of the model explanation detailing our additions to the model.
+*To answer your previous comment we introduced a subsection of the model explanation detailing our additions to the model.
 The Updating model parameter estimates section of the methods contains a discussion of differences between the original and updated analysis.
@@
 As you say, parameter uncertainty for incidence rates could have been included at the time but was not, whilst the updated value for the serial interval has only recently become available.
-This is now explicitly stated in the rewored methods paragraph.*
+This is now explicitly stated in the reworded methods paragraph.*
 
@@
 It highlights the importance of including both parameter and measurement error, as excluding these sources of variation may lead to spuriously precise results.
-This is problematic for policy makers as the worst case scenario often needs to be considered.
-In addition, my exploration of the assumptions used to estimate the annual change in TB incidence rates in the ethnic White UK born illustrates the structural impact of assuming an annual decrease in TB incidence rates.
+This is problematic for policy makers as the worst case scenario often needs to be considered when making policy decisions.
+In addition, our exploration of the assumptions used to estimate the annual change in TB incidence rates in the ethnic White UK born illustrates the structural impact of assuming an annual decrease in TB incidence rates.
 More realistic estimates of the annual decrease in incidence rates resulted in a greatly increased impact of ending the BCG schools scheme.
@@
 
-*Thank you - corrected throughtout.*
+*Thank you - corrected throughout.*
 
@@
 
-*"The simulation approach used here is not the most accurate method for estimating the impact of ending the BCG schools’ scheme as it relies on numerous assumptions based on the available knowledge in 1987.
-The strength of this work is that the estimates were based on the framework used to inform policy making.
+*"The simulation approach used here is not the most accurate method for estimating the impact of ending the BCG schools scheme as it relies on numerous assumptions based on the available knowledge in 1987 and does not account for the role of non-White and non-UK born cases.
+However, the strength of this work is that the estimates were based on the framework used to inform policy making.
 This allowed the strength of model used in the decision-making process to be assessed once parameter uncertainty had been incorperated.
-This would not have been possible if the impact had been assessed using only the observed data.
+This would not have been possible if the impact had been assessed using only the observed data or with a more complex model.
 It also allowed estimates based on updated data to be compared to historic estimates within the same framework.
-If a different framework had been used then these estimates would not have been comparable."*
+This would also not have been possible if a different framework had been used."*
 
@@
 
-*Unfortunately, our choice of "-" to seperate estimates the 2.5% and 97.5% quantile made understanding this paragraph difficult.
-We have moved to using "," as this more clearly identifies when estimates are negative.
+*Unfortunately, our choice of "-" to seperate estimates of the 2.5% and 97.5% quantiles made understanding this paragraph difficult.
+We have moved to using "," throughout the paper as this more clearly identifies when estimates are negative.
 Thank you for picking this up.*
@@
 Sutherland et al.
-assumed a year-on-year decline of 9% with an estimated incidence of around 23 per 100,000 in 1983 for the unvaccinated population.
-In 2013 this would give an incidence rate of around 2.6.
+assumed a year-on-year decline of 9% with an estimated incidence of  ~ 23 per 100,000 in 1983 for the unvaccinated population.
+In 2013 this would give an incidence rate of ~ 2.6.
 These very small incidence rates mean that the difference between incidence rates in unvaccinated and vaccinated cases is also very small leading to the spuriously high numbers need to vaccinate.
-In comparison a 2% decline would give an inidence rate of 12.5.
+In comparison a 2% decline would give an inidence rate of ~ 12.5.
 Our additions to the methods section add extra detail on how these estimates are arrived at.*

### 2019-09-05 peer-review/review-response.Rmd
abstract updated

@@
 
-Thank you for sending review reports for our manuscript. The reviewer’s comments were useful and have improved the paper.
+Thank you for sending review reports for our manuscript. The reviewer’s comments were useful and have greatly improved the paper.
 
@@
 
-**Abstract updates still a work in progress.**
+*Thank you for these helpful insights - we agree that the manuscript needed revisions.
+We have updated the abstract based on your comments (see the marked up paper) and some of the other changes to the paper.
+Unfortunately, due to word limit constraints it was not possible to fully detail how the model was updated (i.e the specifics) at the same time as detailing the results and conclusions.*
 
@@
 *We agree that without the context of the SI and the Sutherland et al.
-paper the methods section was difficult to follow.
-We have greatly expanded on the explanation of the model given in the methods incorperating details from the SI and improved the discussion of the Sutherland et al.
+paper that the methods section was difficult to follow.
+We have greatly expanded on the explanation of the model given in the methods, incorperating details from the SI, and improved the discussion of the Sutherland et al.
 model.*
@@
 
-*We have expanded on this in the introduction as follows, *
-
-*"In England and Wales, universal school-aged (at 13-14 years old) vaccination was introduced after a MRC trial in the 1950s estimated BCG’s effectiveness at 78% in the ethnic White UK born population.[24] The policy remained in place until 2005, when England and Wales changed to targeted vaccination of high risk neonates.
-The 2005 change in BCG vaccination policy was motivated by evidence of decreased transmission of TB, an increasing proportion of TB cases occurring in the non-UK born,[1] and modelling evidence that suggested stopping the BCG schools scheme would have minimal long term effects on incidence rates.[39]"*
+*We have added the age at vaccination (13-14 years old) to the introduction and also discussed vaccination policy post 2005.
+We have also moved a detailed description of the Sutherland et al.
+model from the SI into the main paper which covers coverage assumptions etc.*
 
@@
 *We apologise for the confusion - hopefully our updates to the methods section has helped clarify the model structure.
-We considered the annual decrease in a scenario analysis but all other parameters were otherwise updated.*
+We considered the annual decrease in a scenario analysis but all other parameters were otherwise updated.
+It is variation in these parameters that we are referring to here.*

### 2019-09-05 peer-review/review-response.Rmd
review comments first draft

@@
 
-**What do we think the study design is here?**
+*We have updated the title to,*
+
+*"Reassessing the evidence for universal school-age Bacillus Calmette Guerin (BCG) vaccination in England and Wales: re-evaluating and updating a modelling study"*
 
@@
 
-**I'm not sure this checklist is really appropriate for this paper. It is primarly designed for stastical models used for prediction.**
+*We feel that this checklist is not completely appropriate for this study as it is more focussed towards statistical models designed to be used as tools for prediction.
+However, we would be happy to fill in this or any other checklist if required.*
 
@@
 *To answer your previous comment we introduced a subsection of the model explanation detailing our additions to the model.
-The Updating model parameter estimates section of the methods contains a discussion of differences between the original and updated analysis.
+The Updating model parameter estimates and incorporating parameter uncertainty section of the methods contains a discussion of differences between the original and updated analysis.
 We have reworded this for clarity (please see the marked up paper).

### 2019-09-06 peer-review/review-response.Rmd
finalised review changes

@@
 
-*"We did not involve patients or the public in design or planning of this study."*
+*"We did not involve patients or the public in the design or planning of this study."*
 
@@
 
-*We feel that this checklist is not completely appropriate for this study as it is more focussed towards statistical models designed to be used as tools for prediction.
+*We feel that the TRIPOD  checklist is not completely appropriate for our study because we use an infectious disease model rather than a statistical model designed to be used as a tool for prediction, thus several of the checklist items are not relevant.
 However, we would be happy to fill in this or any other checklist if required.*
@@
 
-*The reviewer is correct in identfying this as an issue.
-We have included some of this uncertainty by re-estimating incidence rates for both those BCG vaccinated and unvaccinated (this is uncertainty is then passed through to the efficacy estimates).
+*We agree with the reviewer that this is an important issue.
+We have included some of this uncertainty by re-estimating incidence rates for both those BCG vaccinated and unvaccinated (this uncertainty is then passed through to the efficacy estimates).
 This approach makes use of the available data on the effiacy of the BCG vaccine in England and Wales.
@@
 
-*We agree. We have reformulated these equations for greater clarity (please see the marked up paper).*
+*We have reformulated these equations for greater clarity (please see the marked up paper).*
 
@@
 
-*Thank you - corrected.*
+*Yes, thank you - corrected.*
 
@@
 
-*The range of uncertainty here is within the scenario. This is due to the incorperation of uncertainty detailed in the Updating model parameter estimates section of the methods.*
+*The range of uncertainty here is within the scenario.
+This is due to the incorperation of uncertainty detailed in the Updating model parameter estimates section of the methods.
+We have reworded this section for clarity.*
 
@@
 *We agree that including transmission structure is important when assessing the impact of infectious disease programmes.
-We did not consider seperating these estimates as we felt it was well established in the literature that transmission structure should always be included.
+We did not consider separating these estimates as we felt it was well established in the literature that transmission structure should always be included.
 Whilst adding to this body of evidence would be worthwhile we feel that adding additional outcome measures may further complicate the results section of this paper.*
@@
 
-*Thank you for these helpful insights - we agree that the manuscript needed revisions.
-We have updated the abstract based on your comments (see the marked up paper) and some of the other changes to the paper.
-Unfortunately, due to word limit constraints it was not possible to fully detail how the model was updated (i.e the specifics) at the same time as detailing the results and conclusions.*
+*Thank you for these helpful insights - we agree that the abstract needed revisions.
+We have updated it based on your comments (see the marked up paper) and some of the other changes to the paper.
+Unfortunately, due to word limit constraints it was not possible to fully detail how the model was updated at the same time as detailing the results and conclusions.*
 
@@
 However, the strength of this work is that the estimates were based on the framework used to inform policy making.
-This allowed the strength of model used in the decision-making process to be assessed once parameter uncertainty had been incorperated.
+This allowed the strength of model used in the decision-making process to be assessed once parameter uncertainty had been incorporated.
 This would not have been possible if the impact had been assessed using only the observed data or with a more complex model.
@@
 
-*We discuss alternative approaches for assessing the impact in the final paragraph of the discussion but you are correct in saying that the rationale of this work was to assess the impact of updateingree an existing framework.
+*We discuss alternative approaches for assessing the impact in the final paragraph of the discussion but you are correct in saying that the rationale of this work was to assess the impact of updating an existing framework.
 We have clarified this by adding the following sentence to the introduction,*
@@
 
-*Unfortunately, our choice of "-" to seperate estimates of the 2.5% and 97.5% quantiles made understanding this paragraph difficult.
+*Unfortunately, our choice of "-" to separate estimates of the 2.5% and 97.5% quantiles made understanding this paragraph difficult.
 We have moved to using "," throughout the paper as this more clearly identifies when estimates are negative.
@@
 In 2013 this would give an incidence rate of ~ 2.6.
-These very small incidence rates mean that the difference between incidence rates in unvaccinated and vaccinated cases is also very small leading to the spuriously high numbers need to vaccinate.
+These very small incidence rates mean that the difference between incidence rates in unvaccinated and vaccinated cases is also very small leading to the spuriously high number need to vaccinate.
 In comparison a 2% decline would give an inidence rate of ~ 12.5.
@@
 *We agree that discounting should be included for economic evaluation.
-However we have not seen it done widely when the study is focussing purely on the epidemiological impact of vaccination.
+However we have not seen it done widely when the study is focusing purely on the epidemiological impact of vaccination.
 We have added the following sentences to the discussion,*
@@
 
-*We agree that this study does not fully answer whether or not the policy change was the correct decision and agree that analysising recent data may be a more valuable approach for this (see here for a preprint by 3 of this studies authors doing this: https://doi.org/10.1101/567511).
+*We agree that this study does not fully answer whether or not the policy change was the correct decision and agree that analysing recent data may be a more valuable approach for this (see here for a preprint by the authors doing this: https://doi.org/10.1101/567511).
 The aim of this study was instead to assess the strength of the evidence used for decision making at the time of the policy change - we have updated the abstract and introduction to clarify this.
 As mentioned by the second reviewer models are rarely evaluated once the policy decision has been made.
-In this study we show that the model used for deicision making was spuriously precise and made several assumptions that led to what we now know were entirely unrealistic estimates.
-These findings are important for decision makers to understand when evaluating quantative evidence for future policy decisions.*
+In this study we show that the model used for decision making was spuriously precise and made several assumptions that led to what we now know were entirely unrealistic estimates.
+These findings are important for decision makers to understand when evaluating quantitative evidence for future policy decisions.*
 
@@
 
-*We agree that without the context of the SI and the Sutherland et al.
-paper that the methods section was difficult to follow.
-We have greatly expanded on the explanation of the model given in the methods, incorperating details from the SI, and improved the discussion of the Sutherland et al.
-model.*
+*We have greatly expanded on the explanation of the model given in the methods, incorperating details from the SI, and improved the discussion of the Sutherland et al. model.*
 
@@
 
-*We have incorperated details from the SI in order to expand on this - please see the marked up paper.*
+*We have incorporated details from the SI in order to expand on this - please see the marked up paper.*
 
@@
 Data on incidence rates in the White UK born was not available.
-Therefore, we used two proxy measures notifications in england and Wales and incidence rates in the UK born in England only (data for Wales was not available).
+Therefore, we used two proxy measures notifications in England and Wales and incidence rates in the UK born in England only (data for Wales was not available).
 Based on your comments (and those of the other reviewers) we have reworded this paragraph.*

### 2019-09-06 README.Rmd
finalised review changes

@@
 
-## Objectives
+## Abstract
+
+### Objectives
 
 In 2005, England and Wales switched from universal BCG vaccination against tuberculosis (TB) disease for school-age children to targeted vaccination of neonates.
-We aimed to assess the quantitative evidence that informed this policy change.
+We aimed to recreate and re-evaluate a previously published model, the results of which informed this policy change.
 
-## Design
+### Design
 
-We recreated a previous approach for estimating the impact of ending the BCG schools’ scheme, updating the model with parameter uncertainty.
-We investigated scenarios considered by the UK’s Joint Committee on Vaccination and Immunisation, and explored new approaches using notification data.
+We recreated an approach for estimating the impact of ending the BCG schools scheme, updating the model with parameter uncertainty and improved parameter estimates where possible.
+We investigated scenarios for the assumed annual decrease in TB incidence rates considered by the UK’s Joint Committee on Vaccination and Immunisation and explored alternative scenarios using notification data.
 
-## Setting
+### Setting
 
@@
 
-## Outcome measures
+### Outcome measures
 
-The number of vaccines needed to prevent a single notification, and the average annual additional notifications caused by ending the BCG schools’ scheme.
+The number of vaccines needed to prevent a single notification, and the average annual additional notifications caused by ending the policy change.
 
-## Results
+### Results
 
-We found a 1.9% annual decrease in TB incidence rates best matched notification data.
-We estimate that 1600 (2.5-97.5% Quantiles (Q): 1300-2100) vaccines would have been required to prevent a single notification in 2004.
-If the scheme had ended in 2001, 302 (2.5-97.5% Q: 238-369) additional annual notifications would have occurred compared to if the scheme had continued.
-If the scheme ended in 2016, 120 (2.5-97.5% Q: 88-155) additional annual notifications would have occurred.
+The model produced predictions with wide confidence intervals when parameter uncertainty was included.
+Estimates using the original assumptions greatly underpredicted the impact of ending school-age vaccination compared to estimates using notification data.
+Updating the assumed annual decrease in TB incidence rates resulted in estimates that were more comparable with notification data with an annual decrease of 1.9% representing the closest estimate to the data.
+Using this assumption, we estimate that 1600 (2.5,97.5% Quantiles (Q): 1300, 2100) vaccines would have been required to prevent a single notification in 2004.
 
-## Conclusions
+### Conclusions
 
-Our estimates of the impact of ending the BCG schools’ scheme were highly sensitive to the annual decrease in incidence rates.
-The impact of ending the BCG schools’ scheme was found to be greater than previously thought when parameter values were updated, and notification data were used.
-Our results highlight the importance of including uncertainty when forecasting the impact of changes in vaccination policy.
+The previously published estimates of the impact of ending the BCG schools scheme were found to be spuriously precise once updated with parameter uncertainty.
+Our estimates of the impact of ending school-age vaccination were highly sensitive to the annual decrease in incidence rates.
+The impact of ending the BCG schools’ scheme was found to be greater than previously thought when notification data were used.
+Our results highlight the importance of including uncertainty and evaluating multiple scenarios when forecasting the impact of changes in vaccination policy.
 

### 2019-09-06 peer-review/review-response.Rmd
resubmission to BMJ open

@@
 Policy makers should consider these updated estimates when assessing the role of BCG vaccination in those at school-age.
-However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiolgical impact of vaccination only."*
+However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiological impact of vaccination only."*
 
@@
 *"Policy makers should consider these updated estimates when assessing the role of BCG vaccination in those at school-age.
-However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiolgical impact of vaccination only."*
+However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiological impact of vaccination only."*
 

### 2019-09-17 R/sutherland_model.R
added updated chain based model

@@ -16,6 +16,7 @@
 #' \code{\link[AssessBCGPolicyChange]{sutherland_trans_params}}.
 #' @param Sym.Lag Numeric, the generation time between infection and symtoms. Must be smaller or equal to the \code{Cohort.Length}.
 #' Defaults to \code{\link[AssessBCGPolicyChange]{sutherland_gen_time}}.
+#' @param update_chains Logical, defaults to \code{FALSE}. Should the transmission chain model be updated or used as found in Sutherland et al.
 #' @return A list of tables reproducing the results presented in Sutherland et al. The final table estimates the total number of additional
 #' cases from ending the shcheme. In order to provide a complete estimate each 5 year estimate has been multiplied by the cohort length.
 #' @export
@@ -27,14 +28,15 @@
 sutherland_model <- function(Data = sutherland_data, incidence_rates = sutherland_incidence_rates, Rates.Per = 100000,
                              Cohort.Length = 5, Data.start = 1969,
                              Annual.TB.Decrease.Yearly = TB_decrease_as_matrix(sutherland_TB_decrease),
-                             Percentage.Year.One = NULL, trans_params = sutherland_trans_params(), Sym.Lag = 2){
+                             Percentage.Year.One = NULL, trans_params = sutherland_trans_params(), Sym.Lag = 2,
+                             update_chains = FALSE){
 
   ## Define variables
   Cohort <- Data[["Cohort"]]
   Cohort.Ranges <- Data[["Cohort.Ranges"]]
   sim_year <- as.numeric(rownames(Annual.TB.Decrease.Yearly))
   Projected.Cohort <- (max(sim_year) - min(sim_year) + 1) / Cohort.Length
-  trans_params <- sutherland_trans_params(Annual.TB.Decrease.Yearly, Sym.Lag = Sym.Lag)
+  trans_params <- sutherland_trans_params(Annual.TB.Decrease.Yearly, Sym.Lag = Sym.Lag, update_chains = update_chains)
   Expected.Total.Sec.Note <- trans_params$Expected.Total.Sec.Note
   Size.First.Gen <- trans_params$Size.First.Gen
   Avg.Int.All.Sec <- trans_params$Avg.Int.All.Sec
@@ -235,7 +237,7 @@ sutherland_model <- function(Data = sutherland_data, incidence_rates = sutherlan
[…]

### 2019-09-17 R/sutherland_trans_params.R

@@ -4,33 +4,48 @@
 #' assumption for estimating the basic reproduction number.
 #' @param Repro.No If \code{Sutherland} is \code{FALSE} sets the basic reproduction number used to estimate
 #' the transmission chain parameters.
+#' @param update_chains Logical, defaults to \code{FALSE}. Should the transmission chain model be updated or used as found in Sutherland et al.
 #' @return A dlist containing the expected total secondary notifications (Expected.Total.Sec.Note),
 #' the size of the first generation (Size.First.Gen) and the average interval to all secondary cases (Avg.Int.All.Sec).
 #' @export
 #' @inheritParams sutherland_model
 #' @examples
 #'
+#'# Orginal model
 #'sutherland_trans_params()
 #'
+#'# Updated model
+#'sutherland_trans_params(update_chains = TRUE)
 sutherland_trans_params <- function(Annual.TB.Decrease.Yearly = TB_decrease_as_matrix(sutherland_TB_decrease), Sym.Lag = sutherland_gen_time,
                                     Sutherland = TRUE, Repro.No = NULL,
-                                    Cohort.Length = 5)
+                                    Cohort.Length = 5, update_chains = FALSE)
 {
 
-  ################################## Calculate parameters for secondary notifcations model
-  ################ Expected Total no. notifications
-  if (Sutherland) {
-    Expected.Total.Sec.Note <- (1 - rowMeans(Annual.TB.Decrease.Yearly))^Sym.Lag
-  }else {
-    Expected.Total.Sec.Note <- rep(Repro.No, nrow(Annual.TB.Decrease.Yearly))
-    names(Expected.Total.Sec.Note) <- rownames(Annual.TB.Decrease.Yearly)
+  if (!update_chains) {
+    ################################## Calculate parameters for secondary notifcations model
+    ################ Expected Total no. notifications
+    if (Sutherland) {
+      Expected.Total.Sec.Note <- (1 - rowMeans(Annual.TB.Decrease.Yearly))^Sym.Lag
+    }else {
+      Expected.Total.Sec.Note <- rep(Repro.No, nrow(Annual.TB.Decrease.Yearly))
+      names(Expected.Total.Sec.Note) <- rownames(Annual.TB.Decrease.Yearly)
+    }
+
+    ############### Size of first Gen
+    Size.First.Gen <- Expected.Total.Sec.Note / (Expected.Total.Sec.Note + 1)
[…]

### 2019-09-17 R/gen_scenario_results.R
first draft of reviewer updates

@@ -3,7 +3,7 @@
 #' @param scenarios A dataframe of scenarios, see the examples for a structural example.
 #' @param parameter_samples A dataframe of parameter samples, see the examples for an example.
 #' @param cores The number of cores to use, defaults to one
-#'
+#' @param ... Additional arguments passed to \code{sutherland_model}.
 #' @return A dataframe combining the scenarios, parameter samples, and model results
 #' @export
 #' @import magrittr
@@ -20,7 +20,7 @@
 #' population_data = sutherland_data, samples = 10)
 #'
 #' gen_scenario_results(scenarios, parameter_samples, cores = 1)
-gen_scenario_results <- function(scenarios, parameter_samples, cores = 1) {
+gen_scenario_results <- function(scenarios, parameter_samples, cores = 1, ...) {
 
   if (max(scenarios$sample) != max(parameter_samples$sample)) {
     stop("The number of samples in scenarios and parameter_samples must match")
@@ -35,7 +35,7 @@ gen_scenario_results <- function(scenarios, parameter_samples, cores = 1) {
                                                           list_incidence_rates = scenarios$incidence_rates,
                                                           list_percentage_year_one = scenarios$per_year_one,
                                                           list_sym_lag = scenarios$sym_lag,
-                                                          cores = cores)
+                                                          cores = cores, ...)
 
   model_sims <- tibble(output =  map(names(model_sims),
                                      ~rep(., length(flatten(model_sims))/length(names(model_sims)))) %>%

### 2019-09-17 R/sutherland_trans_params.R

@@ -40,7 +40,6 @@ sutherland_trans_params <- function(Annual.TB.Decrease.Yearly = TB_decrease_as_m
     # Define the size of the first generation ---------------------------------
     Size.First.Gen <- (1 - rowMeans(Annual.TB.Decrease.Yearly)) ^ Sym.Lag
 
-
     # Define expected total notifications -------------------------------------
     Expected.Total.Sec.Note <- Size.First.Gen / (1 - Size.First.Gen)
 

### 2019-10-03 R/sutherland_model.R
updated model to deal with review flagged Sutherland issue

@@ -18,7 +18,8 @@
 #' Defaults to \code{\link[AssessBCGPolicyChange]{sutherland_gen_time}}.
 #' @param update_chains Logical, defaults to \code{FALSE}. Should the transmission chain model be updated or used as found in Sutherland et al.
 #' @return A list of tables reproducing the results presented in Sutherland et al. The final table estimates the total number of additional
-#' cases from ending the shcheme. In order to provide a complete estimate each 5 year estimate has been multiplied by the cohort length.
+#' cases from ending the shcheme. In order to provide a complete estimate each 5 year estimate has been multiplied by the cohort length. In addition the
+#' primary additional impacts and total secondary notifications arising from each year are also output based on paper revisions.
 #' @export
 #' @importFrom matrixStats colProds
 #' @examples
@@ -279,7 +280,6 @@ sutherland_model <- function(Data = sutherland_data, incidence_rates = sutherlan
   Round.Total.Note.Year.Inelg <- Total.Note.Subpop(Est.Inelg.Risk,Proj.No.Inelg, Length)
 
 
-
   ################### Total Notifications if scheme continues
 
   Total.Note.Year.Full <- Round.Total.Note.Year.Vac.Rec + Round.Total.Note.Year.Vac.Unrec + Round.Total.Note.Year.Tub.Pos + Round.Total.Note.Year.Inelg
@@ -289,7 +289,6 @@ sutherland_model <- function(Data = sutherland_data, incidence_rates = sutherlan
 
   Stan.Year <- names(Total.Note.Year.Full)[1:(length(names(Total.Note.Year.Full)) + (1 - Length))]
   Total.Effects <- Round.Primary.Additional.Note[,Stan.Year] + Round.Secondary.Additional.Note[,Stan.Year] + dummy.Total.Note.Year.Full[,Stan.Year]
-
 
 
   #################### Total additional notifications from ending the scheme in selected years
@@ -302,11 +301,40 @@ sutherland_model <- function(Data = sutherland_data, incidence_rates = sutherlan
   Output <- list(Est.TB.Risk.Round, Est.Note.Prev.Round, Round.Note.Prev.Year, Round.Total.Note.Year,
                 Total.Note.Year.Full, Round.Primary.Additional.Note, Round.Secondary.Additional.Note,p, Total.Secondary.Note.All.Time,
                 Total.Effects, total_additional_notifications, total_add_nots_all_time)
[…]

### 2019-10-04 R/sutherland_model.R
further updates to deal with flaw

@@ -326,8 +326,8 @@ sutherland_model <- function(Data = sutherland_data, incidence_rates = sutherlan
 
    total_secondary_additional <- t(t(Primary.Additional.Note)*Expected.Total.Sec.Note[index]) %>%
     rowSums() %>%
-    round()
-  {. * Cohort.Length}
+    round() %>%
+    {. * Cohort.Length}
 
 
 

### 2019-10-04 peer-review/review-response.Rmd
updated reviewer repsonses

@@
 
-*We agree that including transmission structure is important when assessing the impact of infectious disease programmes.
-We did not consider separating these estimates as we felt it was well established in the literature that transmission structure should always be included.
-Whilst adding to this body of evidence would be worthwhile we feel that adding additional outcome measures may further complicate the results section of this paper.*
+*We have separated the reporting of primary and secondary additional notifications to highlight the role of the transmission model.*
 
@@
 
-*Thank you for your comments. Based on them we have greatly extended our discussion of the methods used and added further clarification where possible. Please see our detailed responses below.*
+*Thank you for your comments - especially the identification of the flaw with the original Sutherland et al.
+model.
+Based on them we have corrected the transmission chain model, extended our discussion of the methods used, and added further clarification where possible.
+Please see our detailed responses below.*
 
@@
 
-*We agree that this needed further explanation. We have greatly extended the explanation using information from the SI and added further clarification where possible.*
+*We agree that this needed further explanation and that the definition used in 1. of Sutherland et al. was incorrect. We have corrected the Sutherland model and extended the model explanation.*
 
@@
 
-*We agree that it is important to be clear which findings are relevant to policy makers and in what context. We have rewored the 4th discussion paragraph as follows in order to provide more context.*
-
-*"This study indicates that some of the evidence used to justify the 2005 change in BCG vaccination policy may have underestimated the impact of ending the scheme.
-It highlights the importance of including both parameter and measurement error, as excluding these sources of variation may lead to spuriously precise results.
-This is problematic for policy makers as the worst case scenario often needs to be considered when making policy decisions.
-In addition, our exploration of the assumptions used to estimate the annual change in TB incidence rates in the ethnic White UK born illustrates the structural impact of assuming an annual decrease in TB incidence rates.
-More realistic estimates of the annual decrease in incidence rates resulted in a greatly increased impact of ending the BCG schools scheme.
-Policy makers should consider these updated estimates when assessing the role of BCG vaccination in those at school-age.
-However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiological impact of vaccination only."*
+*We agree that it is important to be clear which findings are relevant to policy makers and in what context. We have rewored the 4th discussion paragraph to provide more context.*
 
@@
 
-*"The simulation approach used here is not the most accurate method for estimating the impact of ending the BCG schools scheme as it relies on numerous assumptions based on the available knowledge in 1987 and does not account for the role of non-White and non-UK born cases.
+*" The simulation approach used here, although updated where possible, is not the most accurate method for estimating the impact of ending the BCG schools scheme as it relies on numerous assumptions based on the available knowledge in 1987 and does not account for the role of non-White and non-UK born cases.
 However, the strength of this work is that the estimates were based on the framework used to inform policy making.
-This allowed the strength of model used in the decision-making process to be assessed once parameter uncertainty had been incorporated.
-This would not have been possible if the impact had been assessed using only the observed data or with a more complex model.
+This allowed the strength of model used in the decision-making process to be assessed once parameter uncertainty had been incorporated and for flaws in the model to be identified.
+This would not have been possible if the impact had been assessed using only the observed data or with a alternative model.
 It also allowed estimates based on updated data to be compared to historic estimates within the same framework.
@@
 
-*Added detail from the SI into the main body of the methods. Please see the marked up paper.*
+*Added detail to the methods. Please see the marked up paper.*
 
@@
 
-*We have added the following clarification for local regression + reference,*
+*We have added the following clarification for local regression,*
 
-*"LOESS (locally estimated scatterplot smoothing) is a local regression method.[@FoxSnow:2019] It combines multiple regression models in a k-nearest neighbours meta-model.
+*"LOESS (locally estimated scatterplot smoothing) is a local regression method.
+It combines multiple regression models in a k-nearest neighbours meta-model.
 This approach allows nonlinear trends to be fitted to using a series of linear models."*
@@
 
-*Unfortunately, our choice of "-" to separate estimates of the 2.5% and 97.5% quantiles made understanding this paragraph difficult.
+*Unfortunately, our choice of "-" to separate estimates of the 2.5% and 97.5% quantiles made this paragraph unclear.
 We have moved to using "," throughout the paper as this more clearly identifies when estimates are negative.
@@
 As mentioned by the second reviewer models are rarely evaluated once the policy decision has been made.
-In this study we show that the model used for decision making was spuriously precise and made several assumptions that led to what we now know were entirely unrealistic estimates.
+In this study we show that the model used for decision making was methodologically flawed, was spuriously precise, and made several assumptions that led to what we now know were entirely unrealistic estimates.
 These findings are important for decision makers to understand when evaluating quantitative evidence for future policy decisions.*

### 2019-10-07 peer-review/review-response.Rmd
paper review and edits
editing and cuts to word length

@@
 
-*We agree that this needed further explanation and that the definition used in 1. of Sutherland et al. was incorrect. We have corrected the Sutherland model and extended the model explanation.*
+*We agree that this needed further explanation and that the definition used in 1.
+of Sutherland et al.
+was incorrect.
+We have corrected the Sutherland model and extended the model explanation to include this.*
 
@@
 
-*" The simulation approach used here, although updated where possible, is not the most accurate method for estimating the impact of ending the BCG schools scheme as it relies on numerous assumptions based on the available knowledge in 1987 and does not account for the role of non-White and non-UK born cases.
+*"The simulation approach used here, although updated where possible, is not the most accurate method for estimating the impact of ending the BCG schools scheme as it relies on numerous assumptions based on the available knowledge in 1987 and does not account for the role of non-White and non-UK born cases.
 However, the strength of this work is that the estimates were based on the framework used to inform policy making.

### 2019-10-07 README.Rmd
editing and cuts to word length

@@
 
-## Abstract
-
-### Objectives
+## Objectives
 
@@
 
-### Design
+## Design
 
-We recreated an approach for estimating the impact of ending the BCG schools scheme, updating the model with parameter uncertainty and improved parameter estimates where possible.
+We recreated an approach for estimating the impact of ending the BCG schools scheme, correcting a methodological flaw in the model, updating the model with parameter uncertainty, and improving parameter estimates where possible.
 We investigated scenarios for the assumed annual decrease in TB incidence rates considered by the UK’s Joint Committee on Vaccination and Immunisation and explored alternative scenarios using notification data.
 
-### Setting
+## Setting
 
@@
 
-### Outcome measures
+## Outcome measures
 
@@
 
-### Results
+## Results
 
+The original model greatly underestimated the impact of ending school-age vaccination compared to our updated, corrected, model.
 The model produced predictions with wide confidence intervals when parameter uncertainty was included.
-Estimates using the original assumptions greatly underpredicted the impact of ending school-age vaccination compared to estimates using notification data.
-Updating the assumed annual decrease in TB incidence rates resulted in estimates that were more comparable with notification data with an annual decrease of 1.9% representing the closest estimate to the data.
-Using this assumption, we estimate that 1600 (2.5,97.5% Quantiles (Q): 1300, 2100) vaccines would have been required to prevent a single notification in 2004.
+Estimates were most comparable to notification data when an annual decrease in TB incidence rates of 1.9% was used.
+Using this assumption, we estimate that 1600 (2.5,97.5% Quantiles (Q): 1300, 2000) vaccines would have been required to prevent a single notification in 2004.
 
-### Conclusions
+## Conclusions
 
-The previously published estimates of the impact of ending the BCG schools scheme were found to be spuriously precise once updated with parameter uncertainty.
-Our estimates of the impact of ending school-age vaccination were highly sensitive to the annual decrease in incidence rates.
+The previously published model was found to contain a methodological flaw which, once corrected, resulted in a large increase in the estimated impact of ending the BCG schools scheme.
+The original approach was also found to be spuriously precise once updated with parameter uncertainty.
 The impact of ending the BCG schools’ scheme was found to be greater than previously thought when notification data were used.
-Our results highlight the importance of including uncertainty and evaluating multiple scenarios when forecasting the impact of changes in vaccination policy.
+Our results highlight the importance of independent evaluations of modelling evidence, including uncertainty, and evaluating multiple scenarios when forecasting the impact of changes in vaccination policy.
 

### 2019-10-09 peer-review/review-response.Rmd
changes to reviewer responses based on Hannah feedback

@@
 *We agree with the reviewer that this is an important issue.
-We have included some of this uncertainty by re-estimating incidence rates for both those BCG vaccinated and unvaccinated (this uncertainty is then passed through to the efficacy estimates).
+We have included some of this uncertainty by re-estimating incidence rates for both those BCG vaccinated and unvaccinated.
+These rates are then used to estimate vaccine efficacy.
 This approach makes use of the available data on the effiacy of the BCG vaccine in England and Wales.
@@
 
-*We have extensively reworded and extended this section (please see the marked up version). We have added the following sentence for clarity,*
+*We have extensively reworded and extended this section (please see the marked up version). As part of ths we have added the following sentence for clarity,*
 
@@
 
-*We have reformulated these equations for greater clarity (please see the marked up paper).*
+*Yes the distribution of cases is estimated by age group.
+We have reformulated these equations for greater clarity (please see the marked up paper).
+Note that due to paper changes this section of the model has been moved to the SI.*
 
@@
 We therefore estimated incidence rates between 1984 and 1999, and for the years between notifications surveys (1974-1977 and 1979-1982), using locally estimated scatterplot smoothing (LEOSS) regression fitted to incidence rates published in [11] and the estimated incidence rates from 2000 on-wards.
-LOESS is a local regression method that combines multiple regression models in a k-nearest neighbours meta-model.[@FoxSnow:2019] This approach allows nonlinear trends to be fitted to using a series of linear models.
-For years prior to 1973 the annual decreases were assumed to be the mean of the previous 3 years of data.
+LOESS is a local regression method that combines multiple regression models in a k-nearest neighbours meta-model.[@FoxSnow:2019] This approach allows nonlinear trends to be fitted using a series of linear models.
+For years prior to 1973 the annual decreases were assumed to be the mean of the annual decreases from the previous 3 years.
 For both proxy measures the annual decreases in incidence rates post 2016 were assumed to be the average of the estimates in 2013, 2014, and 2015."*
@@
 
-*To answer your previous comment we introduced a subsection of the model explanation detailing our additions to the model.
+*We have introduced a subsection of the model explanation detailing our additions to the model.
 The Updating model parameter estimates and incorporating parameter uncertainty section of the methods contains a discussion of differences between the original and updated analysis.
@@
 
-*We agree that it is important to be clear which findings are relevant to policy makers and in what context. We have rewored the 4th discussion paragraph to provide more context.*
+*We agree that it is important to be clear which findings are relevant to policy makers and in what context. We have reworded the 4th discussion paragraph to provide more context.*
 
@@
 *"The simulation approach used here, although updated where possible, is not the most accurate method for estimating the impact of ending the BCG schools scheme as it relies on numerous assumptions based on the available knowledge in 1987 and does not account for the role of non-White and non-UK born cases.
-However, the strength of this work is that the estimates were based on the framework used to inform policy making.
-This allowed the strength of model used in the decision-making process to be assessed once parameter uncertainty had been incorporated and for flaws in the model to be identified.
+However, the strength of this work is that the estimates are based on the framework used to inform policy making.
+This allowed the strength of the model used in the decision-making process to be assessed once parameter uncertainty had been incorporated and for flaws in the model to be identified.
 This would not have been possible if the impact had been assessed using only the observed data or with a alternative model.
@@
 Whilst there has been a systematic review the authors identifed only two papers with estimates (the one used here and one other older paper).
-Both papers made use of similar datasets in the Netherlands.
-The cited paper is 10 years newer, has an improved sample size, and improved methods.*
+Both papers made use of similar datasets in the Netherlands, however, the cited paper is most recent (by 10 years) and has an improved sample size and improved methods compared to the older paper.*
 
@@
 
-*We have reworded this paragraph (and made changes suggested by reviewer 1).
+*We have reworded this paragraph (and made additional changes suggested by reviewer 1).
 Please see the marked up paper.
@@
 It combines multiple regression models in a k-nearest neighbours meta-model.
-This approach allows nonlinear trends to be fitted to using a series of linear models."*
+This approach allows nonlinear trends to be fitted using a series of linear models."*
 
@@
 *"Policy makers should consider these updated estimates when assessing the role of BCG vaccination in those at school-age.
-However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiological impact of vaccination only."*
+However, decisions regarding vaccine policy in the UK require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiological impact of vaccination only."*
 

### 2019-10-10 peer-review/review-response.Rmd
updated reviewer reponses based on Ellen feedback

@@
 *We feel that the TRIPOD  checklist is not completely appropriate for our study because we use an infectious disease model rather than a statistical model designed to be used as a tool for prediction, thus several of the checklist items are not relevant.
-However, we would be happy to fill in this or any other checklist if required.*
+However, we would be happy to fill in this checklist if required.*
 
@@
 
-*We agree with the reviewer that this is an important issue.
+*We agree with the reviewer that this is a potentially important issue.
 We have included some of this uncertainty by re-estimating incidence rates for both those BCG vaccinated and unvaccinated.
@@
 
-*Yes the distribution of cases is estimated by age group.
-We have reformulated these equations for greater clarity (please see the marked up paper).
-Note that due to paper changes this section of the model has been moved to the SI.*
+*No.
+The distribution of secondary cases is not differentiated by age.
+In revising the manuscript, we identified an error in the original analysis, which we have corrected in our paper.
+This error means that this section of the model is no longer valid once the model has been updated and so it is only used to validate the original analysis.
+Based on this we have moved this section to the SI.
+However, we have expanded on the model explanation both here and in other sections.*
 

### 2019-10-11 README.Rmd

@@
 The model produced predictions with wide confidence intervals when parameter uncertainty was included.
-Estimates were most comparable to notification data when an annual decrease in TB incidence rates of 1.9% was used.
+Model estimates were most comparable to notification data when an annual decrease in TB incidence rates of 1.9% was used.
 Using this assumption, we estimate that 1600 (2.5,97.5% Quantiles (Q): 1300, 2000) vaccines would have been required to prevent a single notification in 2004.
@@
 The original approach was also found to be spuriously precise once updated with parameter uncertainty.
-The impact of ending the BCG schools’ scheme was found to be greater than previously thought when notification data were used.
+The impact of ending the BCG schools scheme was found to be greater than previously thought when notification data were used.
 Our results highlight the importance of independent evaluations of modelling evidence, including uncertainty, and evaluating multiple scenarios when forecasting the impact of changes in vaccination policy.

### 2019-12-11 peer-review/review-response-2.Rmd
outlined initial review responses

---
title: "Reviewer Response"
output:
  word_document: default
  html_document:
    df_print: paged
---

Dear Editor,

Thank you for sending review reports for our manuscript. The reviewer’s comments were again useful and have greatly improved the paper.

We enclose a revised version of the manuscript (clean and with tracked changes), which addresses the reviewer’s comments and suggestions. The reviewer’s comments are shown below with our response to each comment in italics.

Yours sincerely,

Sam Abbott

## Authors' response to Reviewer #1

The manuscript has been improved and most of my comments have been addressed.

*Thank you for taking the time to review the manuscript again.*

I have only a few minor suggestions: 
The abbreviation 'JCVI' is still spelt wrong in many parts of the manuscript, including figure legends. I suggest the authors run a thorough check.

*Thank you for highlighting this mistake - we have thoroughly checked the manuscript and correction misspellings.*

It seems to me that the two method sections - 'estimating notification rates' and 'construction of forward estimates' - are the content of Sutherland et al.'s model. If they are, marking the origin explicitly (potentially in the section titles) may help the audience to compare between the original and updated methods. Or if the original and updated methods are both involved in these two sections, please state the difference more explicitly.

**

In 'Original transmission chain model' section, I found the notations 'T' confusing, as they appear in both steps 1. and 2. but come with different definitions. Although this may be the original way how the Sutherland et al.'s model was presented, do you consider to assign different notations or add subscripts to them for clarification?

**

In 'Model assumptions' of supplementary information, the notation 'T' is used but without further definition. As there are two definitions for 'T' used in the main text (as mentioned in the above point), it will be helpful to define the notation again here or make it clearly linked back to the definition in the main text.

**

## Authors' response to Reviewer #2

OVERALL

I appreciate the effort made by the authors to improve the clarity and respond to my comments. I find the article is now much easier to follow, and should be published after a few minor revisions.

*Thank you for reviewing the updated manuscript. We agree that the last set of revisions substantially imrproved the manuscript.*

SPECIFIC

- Abstract. I find the Conclusions a bit long and containing Results. I would suggest moving the first two sentences of the current Conclusions up into Results to have a punchier summary & implication in the Conclusions. Also: can you be more specific about the meaning of 'most comparable to notification data'?

*As suggested we have moved the last two sentences of the conclusions into the results section.*

- Strengths & limitations bullets. 'our results are conservative' - I don't find it obvious what this means. Which results, and which direction does 'conservative' mean in this case? (Eg result could be referring to the NNV, although I wouldn't be certain which direction conservative would then mean; could be cases caused by change, in which case conservative means this could be larger; or could be talking about the increase in uncertainty.)

- Model description in text. This is now much clearer. There are a few minor suggestions for the section on Updating the transmission chain model.

In this section, 'secondary cases' refers (in introducing z) to those directly resulting in the next generation. In bullet 3 introducing Z, all secondary cases are considered. Suggest using more specific phraseology, eg '(secondary) cases in the next generation' where it applies.

Related to this perhaps, I didn't understand the bit about not being able to 'estimate when secondary notifications occurred' when you have just defined them as z years later, or if they are not the next generation cases you have a distribution implied by bullet point defining Z. If this isn't important and keeping them in the same year as root cases just facilitates book-keeping, then this should come across.

Also around here, referring to x as the 'initial generation size' leaves room for confusion on two fronts: a) initial = zeroth or first. It seems you mean first, but could be understood the other way. b) it is a relative size (to the zeroth generation).

- Model description in SI. While much clearer in the article, I'm still a bit confused by the equations in the start of the SI. Could these be explained? I don't really understand what the first equation is meant to mean: eg what does 'current' mean (is it the notifications in a year with data, to be matched to data?)? How do you distinguish notifications as primary vs secondary (and if defined by model dynamics, how is the initial split determined)? Why are x and f making an appearance? For the second equation, why is a fraction (1-fx) of the current propagated and why is this multiplied by this decline factor? In the first instance, x already captures information about decline. In the second instance, the power L-Z is mysterious to me (I was expecting a power like L/z to count generations, with little z rather than Z).

MINUTIAE

*Thanks for highlighting the JVCI typo along with the other issues outlined below. We have thoroughly checked the manuscript and corrected the remaining misspellings of JVCI. We have also dealt with your other points as suggested.*

- I would note that the typo 'JVCI' picked up by another reviewer still occurs in many places, including Figure captions

- LOESS makes an appearance as LEOSS

- 'DHSS' appears before definition

- "13 year old's"

- 'can not': more usual to have cannot

- In various places (eg line 12 page 6): 'decay of 9% in notification rates' - need to specify annual or per year every time.

### 2019-12-12 peer-review/review-response-2.Rmd
added correction of JCVI
review 1 responses

@@
 
-Thank you for sending review reports for our manuscript. The reviewer’s comments were again useful and have greatly improved the paper.
+Thank you for sending review reports for our manuscript. The reviewer’s comments were again useful.
 
@@
 
-*Thank you for highlighting this mistake - we have thoroughly checked the manuscript and correction misspellings.*
+*Thank you for highlighting this mistake - we have thoroughly checked the manuscript and corrected misspellings.*
 
@@
 
-**
+*We have marked these sections as original (to match the approach used for the transmission model.*
 
@@
 
-**
+*We have standardised the definitions used for T in 1. and 2. *
 
@@
 
-**
+*We have added the standardised definition now used in both 1. and 2. to this section along with the definition for $x$.*
 

### 2019-12-12 peer-review/review-response-2.Rmd
reviewer 2 minor responses

@@
 
-*Thank you for highlighting this mistake - we have thoroughly checked the manuscript and corrected misspellings.*
+*Thank you for highlighting this mistake and apologies that it was not dealt with correctly in the last round of responses - we have thoroughly checked the manuscript and corrected misspellings.*
 

### 2019-12-12 peer-review/review-response-2.Rmd
initial draft of review responses and revisions

@@
 
-Thank you for sending review reports for our manuscript. The reviewer’s comments were again useful.
+Thank you for sending review reports for our manuscript.
+The previous round of reviewer comments greatly improved the manuscript.
+The reviewer comments from this round were again useful and have led to further improvements.
 
@@
 
-*Thank you for taking the time to review the manuscript again.*
+*Thank you for taking the time to review the manuscript again and for your comments.*
 
@@
 
-*Thank you for reviewing the updated manuscript. We agree that the last set of revisions substantially imrproved the manuscript.*
+*Thank you for reviewing the updated manuscript and for your suggested revisions. We agree that the last set of revisions substantially improved the manuscript. *
 
@@
 
-*As suggested we have moved the last two sentences of the conclusions into the results section.*
+*As suggested we have moved the first two sentences of the conclusions into the results section, editing them to avoid repetition.
+We have reworded 'most comparable to notification data' to make it clear that model estimates based on an assumed decrease in TB are being compared to model estimates derived using notification data.
+*
 
@@
 (Eg result could be referring to the NNV, although I wouldn't be certain which direction conservative would then mean; could be cases caused by change, in which case conservative means this could be larger; or could be talking about the increase in uncertainty.)
+
+*We have reworded this sentence (and the corresponding sentence in the discussion) to indicate that we think that excluding these factors may mean our results are underestimates.*
 
@@
 
+*Apologies for the confusion. We have used the more specifc phraseology that you suggested.*
+
 Related to this perhaps, I didn't understand the bit about not being able to 'estimate when secondary notifications occurred' when you have just defined them as z years later, or if they are not the next generation cases you have a distribution implied by bullet point defining Z.
 If this isn't important and keeping them in the same year as root cases just facilitates book-keeping, then this should come across.
+
+*This statement was intended to indicate that our approach could not allocate cases annually (only by generation).
+This step was neccessary in order to completely reproduce the previously published results but does not impact estimates of the overall impact of ending the scheme.
+We have clarified this both in the methods section and in the SI.*
 
@@
 b) it is a relative size (to the zeroth generation).
+
+*We have updated the definition to be the relative generation size as it is a relative size to the previous generation (i.e. generation $i$ is based on generation $i - 1$).*
 
@@
 
+*We agree that this could have been better explained.
+Hopefully, our changes from your above points add some clarity about what is being modelled.
+The aim of this step is to allocate secondary notifications to either the current year being considered or the subsequent modelled time step.
+This was found to be neccessary in order to reproduce Sutherland et al.'s final results table but does not impact the overall model results and so is only used for model validation purposes (as is hopefully now clear from the paper).*
+
 MINUTIAE
@@
 *Thanks for highlighting the JVCI typo along with the other issues outlined below.
-We have thoroughly checked the manuscript and corrected the remaining misspellings of JVCI.
+We have thoroughly checked the manuscript and corrected the remaining misspellings.
 We have also dealt with your other points as suggested.*

### 2019-12-12 README.Rmd
initial draft of review responses and revisions

@@
 
-The original model greatly underestimated the impact of ending school-age vaccination compared to our updated, corrected, model.
+The previously published model was found to contain a methodological flaw and to be spuriously precise.
+It greatly underestimated the impact of ending school-age vaccination compared to our updated, corrected, model.
 The model produced predictions with wide confidence intervals when parameter uncertainty was included.
@@
 
-The previously published model was found to contain a methodological flaw which, once corrected, resulted in a large increase in the estimated impact of ending the BCG schools scheme.
-The original approach was also found to be spuriously precise once updated with parameter uncertainty.
 The impact of ending the BCG schools scheme was found to be greater than previously thought when notification data were used.

### 2019-12-13 README.Rmd
updated to final reviewer revisions

@@
 It greatly underestimated the impact of ending school-age vaccination compared to our updated, corrected, model.
-The model produced predictions with wide confidence intervals when parameter uncertainty was included.
-Model estimates were most comparable to notification data when an annual decrease in TB incidence rates of 1.9% was used.
+The updated model produced predictions with wide confidence intervals when parameter uncertainty was included.
+Model estimates based on an assumption of an annual decrease in TB incidence rates of 1.9% were closest to those estimated using notification data.
 Using this assumption, we estimate that 1600 (2.5,97.5% Quantiles (Q): 1300, 2000) vaccines would have been required to prevent a single notification in 2004.

### 2019-12-13 peer-review/review-response-2.Rmd

@@
 
-*Thank you for reviewing the updated manuscript and for your suggested revisions. We agree that the last set of revisions substantially improved the manuscript. *
+*Thank you for reviewing the updated manuscript and for your suggested revisions. We agree that the last set of revisions substantially improved the manuscript.*
 
@@
 *As suggested we have moved the first two sentences of the conclusions into the results section, editing them to avoid repetition.
-We have reworded 'most comparable to notification data' to make it clear that model estimates based on an assumed decrease in TB are being compared to model estimates derived using notification data.
-*
+We have reworded 'most comparable to notification data' to make it clear that model estimates based on an assumed decrease in TB are being compared to model estimates derived using notification data.*
 
@@
 
-*We have reworded this sentence (and the corresponding sentence in the discussion) to indicate that we think that excluding these factors may mean our results are underestimates.*
+*We have reworded this sentence (and the corresponding sentence in the discussion) to indicate that we think that excluding these factors may mean our results underestimate the impact of ending the BCG schools scheme.*
 
@@
 
-*Apologies for the confusion. We have used the more specifc phraseology that you suggested.*
+*We have updated the text to use the more specifc phraseology suggested.*
 
@@
 
-*This statement was intended to indicate that our approach could not allocate cases annually (only by generation).
+*This statement was intended to indicate that our approach could not allocate cases annually (only by generation or overall).
 This step was neccessary in order to completely reproduce the previously published results but does not impact estimates of the overall impact of ending the scheme.
@@
 Hopefully, our changes from your above points add some clarity about what is being modelled.
-The aim of this step is to allocate secondary notifications to either the current year being considered or the subsequent modelled time step.
-This was found to be neccessary in order to reproduce Sutherland et al.'s final results table but does not impact the overall model results and so is only used for model validation purposes (as is hopefully now clear from the paper).*
+The aim of this step is to allocate secondary notifications to either the current year being considered or the subsequent modelled time step (5 years in the future).
+This was found to be neccessary in order to reproduce Sutherland et al.'s final results table but does not impact the overall model results and so is only used for model validation purposes (as is hopefully now clear from the paper).
+We updated the description of these equations in the SI in order to provide more context.*
 
